---
title: "Lipidomics Normalization with pmartR"
author: "Kaitlin Rempfert"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged 
    number_sections: yes 
    css: stylesheet.css 
    toc: yes 
    toc_float: true 
    toc_depth: 3 
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# load required packages
library(tidyverse)
library(ggplot2)
library(pmartR)
library(pander)
library(emmeans)
library(openxlsx)
```


# Positive Mode

## Read in lipidomic data object in pmartR

```{r}
mylipidData_pos <- readRDS("output/pmartR/mylipidData_pos_batch3.rds")
```

Need to go back and reprocess raw data without corn samples, but for now just filter out 
```{r}
#vector of corn samples
corn <- mylipidData_pos$f_data %>% filter(crop == "Corn")
Corn <- corn$sample_name

myfilt <- custom_filter(mylipidData_pos, f_data_remove = Corn)
mylipidData_pos <- applyFilt(myfilt, mylipidData_pos)
```

Missing values (or 0's) are represented by NA, and data were log2 transformed.
```{r}
#replace 0's with NAs -- this is now done automatically when the data object is created via as.lipidData 
mylipidData_pos <- edata_replace(mylipidData_pos, x = 0, y = NA)

#log transform the data
mylipidData_pos <- edata_transform(mylipidData_pos, data_scale = "log2")
attributes(mylipidData_pos)$data_info$data_scale

# shorten sample names for plotting purposes
mylipidData_pos <- custom_sampnames(mylipidData_pos, delim = "_", components = c(1:4))

# define groups based on the "group" variable
mylipidData_pos <- group_designation(mylipidData_pos, main_effects = "group", batch_id = "batch")

plot(mylipidData_pos, use_VizSampNames = TRUE, color_by = "group", order_by = "group")
```


Quality control (QC) samples were run throughout the batches of this study, however, given previous unsuccessful attempts at batch effect correction, only the data from the more recent batch is being used. Therefore, we have no need of the QC samples going forward and we remove them from the dataset now. 

```{r}
#vector of QC samples
qc <- mylipidData_pos$f_data %>% filter(group == "qc")
QC <- qc$sample_name

myfilt <- custom_filter(mylipidData_pos, f_data_remove = QC)
mylipidData_pos <- applyFilt(myfilt, mylipidData_pos)
```




Below, we look at the number of samples in each site-by-timepoint combination by batch. We are restricting the data to a single batch, numbered "3".

**Table: Group by Batch**
```{r}
pander(table(mylipidData_pos$f_data$group, mylipidData_pos$f_data$batch, useNA = "ifany"))
```




## Quality Control

All lipids were present in at least 2 samples within each batch.

```{r}
myfilter <- molecule_filter(mylipidData_pos)
summary(myfilter, min_num = 2)
# filter not applied because it doesn't remove any lipids
```


```{r}
# look for potential sample outliers
myfilter <- rmd_filter(mylipidData_pos, metrics = c("MAD", "Skewness", "Correlation"))
temp <- summary(myfilter, pvalue_threshold = 0.0001)
potential_outliers <- temp$filtered_samples
```

RMD-PAV is used to identify potential outliers, of which `r length(potential_outliers)` are identified: 

`r print(potential_outliers)`

```{r}
plot(myfilter, pvalue_threshold = 0.0001, use_VizSampNames = TRUE)

if(length(potential_outliers) > 0){
  for(i in 1:length(potential_outliers)){
    print(plot(myfilter, sampleID = potential_outliers[i], use_VizSampNames = TRUE))
  }
}

mylipidData_pos$f_data[which(mylipidData_pos$f_data$sample_name %in% potential_outliers), ]
```


```{r}
mycor <- cor_result(mylipidData_pos)
plot(mycor, interactive = TRUE)

# nothing looks terribly egregious here
```


```{r}
mydimres <- dim_reduction(mylipidData_pos)
plot(mydimres, interactive = TRUE)

mypca <- data.frame(SampleID = mydimres$SampleID, PC1 = mydimres$PC1, PC2 = mydimres$PC2)

```


We remove 1 sample as an outlier: EC5063_L2_Pos_25Jan20.

```{r}
myfilt <- custom_filter(mylipidData_pos, f_data_remove = c("EC5063_L2_Pos_25Jan20_Brandi_WCSH115319.mzXML"))
mylipidData_pos <- applyFilt(myfilt, mylipidData_pos)
```


## Normalization


```{r}
# median center the data
mylipidData_pos <- normalize_global(mylipidData_pos, subset_fn = "all", norm_fn = "median", backtransform = TRUE, apply_norm = TRUE)

plot(mylipidData_pos, use_VizSampNames = TRUE, color_by = "group", order_by = "group")

saveRDS(mylipidData_pos, "output/lipid_pos_median.RDS")
```


Look at PCA plot after normalization:

```{r}
mydimres <- dim_reduction(mylipidData_pos)
plot(mydimres, interactive = TRUE)

mypca <- data.frame(SampleID = mydimres$SampleID, PC1 = mydimres$PC1, PC2 = mydimres$PC2)
```



# Negative Mode

## Read in lipidomic data object in pmartR
```{r}
mylipidData_neg <- readRDS("output/pmartR/mylipidData_neg_batch4.rds")
```

Need to go back and reprocess raw data without corn samples, but for now just filter out 
```{r}
#vector of corn samples
corn <- mylipidData_neg$f_data %>% filter(crop == "Corn")
Corn <- corn$sample_name

myfilt <- custom_filter(mylipidData_neg, f_data_remove = Corn)
mylipidData_neg <- applyFilt(myfilt, mylipidData_neg)
```

Missing values (or 0's) are represented by NA, and data were log2 transformed.
```{r}
#replace 0's with NAs
mylipidData_neg <- edata_replace(mylipidData_neg, x = 0, y = NA)

#log transform the data
mylipidData_neg <- edata_transform(mylipidData_neg, data_scale = "log2")
# attributes(mylipidData_neg)$data_info$data_scale

# shorten sample names for plotting purposes
mylipidData_neg <- custom_sampnames(mylipidData_neg, delim = "_", components = c(1:4))

# define groups babsed on the "group" variable
mylipidData_neg <- group_designation(mylipidData_neg, main_effects = "group", batch_id = "batch")

plot(mylipidData_neg, use_VizSampNames = TRUE, color_by = "group", order_by = "group")
```


Quality control (QC) samples were run throughout the batches of this study, however, given previous unsuccessful attempts at batch effect correction, only the data from the more recent batch is being used. Therefore, we have no need of the QC samples going forward and we remove them from the dataset now. 

```{r}
#vector of QC samples
qc <- mylipidData_neg$f_data %>% filter(group == "qc")
QC <- qc$sample_name

myfilt <- custom_filter(mylipidData_neg, f_data_remove = QC)
mylipidData_neg <- applyFilt(myfilt, mylipidData_neg)
```

Below, we look at the number of samples in each site-by-timepoint combination by batch. We are restricting the data to a single batch, numbered "3".

**Table: Group by Batch**

```{r}
pander(table(mylipidData_neg$f_data$group, mylipidData_neg$f_data$batch, useNA = "ifany"))
```


## Quality Control

All lipids were present in at least 2 samples within each batch.

```{r}
myfilter <- molecule_filter(mylipidData_neg)
summary(myfilter, min_num = 2)
# filter not applied because it doesn't remove any lipids
```

```{r}
# look for potential sample outliers
myfilter <- rmd_filter(mylipidData_neg, metrics = c("MAD", "Skewness", "Correlation"))
temp <- summary(myfilter, pvalue_threshold = 0.0001)
potential_outliers <- temp$filtered_samples
```

RMD-PAV is used to identify potential outliers, of which `r length(potential_outliers)` is/are identified: 

`r print(potential_outliers)`

```{r}
plot(myfilter, pvalue_threshold = 0.0001, use_VizSampNames = TRUE)#, order_by = "group")

if(length(potential_outliers) > 0){
  for(i in 1:length(potential_outliers)){
    print(plot(myfilter, sampleID = potential_outliers[i], use_VizSampNames = TRUE))
  }
}

mylipidData_neg$f_data[which(mylipidData_neg$f_data$sample_name %in% potential_outliers), ]

# 32  EC5063_L2_Neg_02Feb21_Brandi_WCSH115319.mzXML WS_2m 4 249 Wisconsin Switchgrass 2Month
# 32 EC5063 --> was also outlier in positive mode
```

```{r}
mycor <- cor_result(mylipidData_neg)
plot(mycor, interactive = TRUE)
```

```{r}
mydimres <- dim_reduction(mylipidData_neg)
plot(mydimres, interactive = TRUE)

mypca <- data.frame(SampleID = mydimres$SampleID, PC1 = mydimres$PC1, PC2 = mydimres$PC2)
```

We remove 1 sample as an outlier: EC5063_L2_Neg_02Feb21. This is the same sample that was removed from the Positive Mode data.

```{r}
myfilt <- custom_filter(mylipidData_neg, f_data_remove = c("EC5063_L2_Neg_02Feb21_Brandi_WCSH115319.mzXML"))
mylipidData_neg <- applyFilt(myfilt, mylipidData_neg)
```


## Normalization

```{r}
# median center the data
mylipidData_neg <- normalize_global(mylipidData_neg, subset_fn = "all", norm_fn = "median", backtransform = TRUE, apply_norm = TRUE)

plot(mylipidData_neg, use_VizSampNames = TRUE, color_by = "group", order_by = "group")

saveRDS(mylipidData_neg,"output/lipid_neg_median.RDS")
```

Look at PCA plot after normalization:

```{r}
mydimres <- dim_reduction(mylipidData_neg)
plot(mydimres, interactive = TRUE)

mypca <- data.frame(SampleID = mydimres$SampleID, PC1 = mydimres$PC1, PC2 = mydimres$PC2)
```



# Combine Positive and Negative mode data

Now that data have been normalized, we combine the negative and positive ionization mode into a single dataset. Sample names have been truncated so that they are identical between the negative and posisive ionization modes.

```{r}
# make the sample names the same between the two ionization modes

# create the new names
new_neg_names <- unlist(lapply(str_split(mylipidData_neg$f_data$sample_name, "_"), function(x) paste(x[1], x[2], sep = "_")))
new_pos_names <- unlist(lapply(str_split(mylipidData_pos$f_data$sample_name, "_"), function(x) paste(x[1], x[2], sep = "_")))

# check some stuff
all(new_neg_names == new_pos_names) # new names for neg and pos all match
all(mylipidData_neg$f_data$sample_name == names(mylipidData_neg$e_data)[-1]) # order of samples in e_data and f_data is the same within neg mode
all(mylipidData_pos$f_data$sample_name == names(mylipidData_pos$e_data)[-1]) # order of samples in e_data and f_data is the same within pos mode

# set the new names
mylipidData_neg$f_data$sample_name <- new_neg_names
names(mylipidData_neg$e_data)[-1] <- new_neg_names
mylipidData_pos$f_data$sample_name <- new_pos_names
names(mylipidData_pos$e_data)[-1] <- new_pos_names
```

Lipid identifiers have been augmented to include the mode in which it was identified.
```{r}
# de-dupe any duplicate edata identifiers
mylipidData_neg$e_data[,get_edata_cname(mylipidData_neg)] <- paste0("mylipidData_neg_", mylipidData_neg$e_data[,get_edata_cname(mylipidData_neg)])
# de-dupe any duplicate emeta identifiers
mylipidData_neg$e_meta[,get_edata_cname(mylipidData_neg)] <- paste0("mylipidData_neg_", mylipidData_neg$e_meta[,get_edata_cname(mylipidData_neg)])

# combine them
mylipidData <- combine_lipidData(mylipidData_pos, mylipidData_neg, retain_groups = T, retain_filters = T)
plot(mylipidData, order_by = "group", color_by = "group")

# make Crop and Site into factors so I can control the order in the models below
mylipidData$f_data$crop <- factor(mylipidData$f_data$crop, levels = c("Switchgrass", "Corn"))
mylipidData$f_data$site <- factor(mylipidData$f_data$site, levels = c("Michigan", "Wisconsin"))

saveRDS(mylipidData, "output/pmartR/mylipidData.RDS")
```

 
