---
title: "Persistent PLFAs"
author: "Kaitlin Rempfert"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged 
    number_sections: yes 
    css: stylesheet.css 
    toc: yes 
    toc_float: true 
    toc_depth: 3 
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# load required packages
library(tidyverse)
library(ggplot2)
library(pmartR)
library(pander)
library(emmeans)
library(openxlsx)
library(ggpubr)
library(rstatix)
library(Rodin)
library(ComplexHeatmap)
```

## Load Data
```{r}
PLFA_df <- readRDS(file = "output/plfas_w_meta.rds")
```

## Enrichment
```{r}
#filter to compounds that were detected at both 2 months and 1 year 
PLFA_df_enrich_filt <- PLFA_df %>% select(c(1:6, wt.perc.13C.excess)) %>% filter(!is.na(wt.perc.13C.excess)) %>% group_by(id, Fraction) %>% mutate(n = sum(!is.na(wt.perc.13C.excess))) %>% filter(n == 2) %>% ungroup() %>% group_by(site, timepoint, Fraction) %>% mutate(reps = sum(!is.na(wt.perc.13C.excess))) %>% ungroup() %>% filter(reps > 3)

#perform t-test (keep sites separate in case a lipid is persistent in one site and not another)
PLFA_enrich_ttest_sep <- PLFA_df_enrich_filt  %>%
  filter(reps > 3) %>%
  select(c(Fraction, site, timepoint, id, wt.perc.13C.excess)) %>%
  arrange(Fraction, site, timepoint, id) %>%
  group_by(Fraction, site) %>%
  t_test(wt.perc.13C.excess ~ timepoint, paired = TRUE, var.equal = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
#perform t-test again without separating sites 
PLFA_enrich_ttest <- PLFA_df_enrich_filt  %>%
  filter(reps > 3) %>%
  select(c(Fraction, site, timepoint, id, wt.perc.13C.excess)) %>%
  arrange(Fraction, site, timepoint, id) %>%
  group_by(Fraction) %>%
  t_test(wt.perc.13C.excess ~ timepoint, paired = TRUE, var.equal = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
    
#calculate Cohen's d (keep sites separate in case a lipid is persistent in one site and not another)
PLFA_enrich_cohens_sep <- PLFA_df_enrich_filt  %>%
  filter(reps > 3) %>%
  select(c(Fraction, site, timepoint, id, wt.perc.13C.excess)) %>%
  arrange(Fraction, site, timepoint, id) %>%
  group_by(Fraction, site) %>%
  cohens_d(wt.perc.13C.excess ~ timepoint, paired = TRUE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
#recalculate without separating sites
PLFA_enrich_cohens <- PLFA_df_enrich_filt  %>%
  filter(reps > 3) %>%
  select(c(Fraction, site, timepoint, id, wt.perc.13C.excess)) %>%
  arrange(Fraction, site, timepoint, id) %>%
  group_by(Fraction) %>%
  cohens_d(wt.perc.13C.excess ~ timepoint, paired = TRUE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))

#combine t-test and cohen's d outputs
PLFA_enrich_stats_sep <- left_join(PLFA_enrich_ttest_sep, PLFA_enrich_cohens_sep)
PLFA_enrich_stats <- left_join(PLFA_enrich_ttest, PLFA_enrich_cohens)

#reformat output for joining for enrichment data 
PLFA_enrich_stats_rf_sep <- left_join(PLFA_df_enrich_filt, PLFA_enrich_stats_sep %>% select(c(Fraction, site, group1, group2, p, effsize)) %>% dplyr::rename("p_enrich" = "p") %>% dplyr::rename("effsize_enrich" = "effsize"))
PLFA_enrich_stats_rf <- left_join(PLFA_df_enrich_filt, PLFA_enrich_stats %>% select(c(Fraction, group1, group2, p, effsize)) %>% dplyr::rename("p_enrich" = "p") %>% dplyr::rename("effsize_enrich" = "effsize"))
```

## Abundance 
```{r}
#filter to compounds that were detected at both 2 months and 1 year 
PLFA_df_abund_filt <- PLFA_df %>% select(c(1:7)) %>% filter(!is.na(ug.Total.C.per.g.dry.soil)) %>% group_by(id, Fraction) %>% mutate(n = sum(!is.na(ug.Total.C.per.g.dry.soil))) %>% filter(n == 2) %>% ungroup() %>% group_by(site, timepoint, Fraction) %>% mutate(reps = sum(!is.na(ug.Total.C.per.g.dry.soil))) %>% ungroup() %>% filter(reps >3) 

#perform t-test (keep sites separate in case a lipid is persistent in one site and not another)
PLFA_abund_ttest_sep <- PLFA_df_abund_filt  %>%
  filter(reps > 3) %>%
  select(c(Fraction, site, timepoint, id, ug.Total.C.per.g.dry.soil)) %>%
  arrange(Fraction, site, timepoint, id) %>%
  group_by(Fraction, site) %>%
  t_test(ug.Total.C.per.g.dry.soil ~ timepoint, paired = TRUE, var.equal = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
#perform t-test again without separating sites 
PLFA_abund_ttest <- PLFA_df_abund_filt  %>%
  filter(reps > 3) %>%
  select(c(Fraction, site, timepoint, id, ug.Total.C.per.g.dry.soil)) %>%
  arrange(Fraction, site, timepoint, id) %>%
  group_by(Fraction) %>%
  t_test(ug.Total.C.per.g.dry.soil ~ timepoint, paired = TRUE, var.equal = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
    
#calculate Cohen's d (keep sites separate in case a lipid is persistent in one site and not another)
PLFA_abund_cohens_sep <- PLFA_df_abund_filt  %>%
  filter(reps > 3) %>%
  select(c(Fraction, site, timepoint, id, ug.Total.C.per.g.dry.soil)) %>%
  arrange(Fraction, site, timepoint, id) %>%
  group_by(Fraction, site) %>%
  cohens_d(ug.Total.C.per.g.dry.soil ~ timepoint, paired = TRUE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
#recalculate without separating sites
PLFA_abund_cohens <- PLFA_df_abund_filt  %>%
  filter(reps > 3) %>%
  select(c(Fraction, site, timepoint, id, ug.Total.C.per.g.dry.soil)) %>%
  arrange(Fraction, site, timepoint, id) %>%
  group_by(Fraction) %>%
  cohens_d(ug.Total.C.per.g.dry.soil ~ timepoint, paired = TRUE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))

#combine t-test and cohen's d outputs
PLFA_abund_stats_sep <- left_join(PLFA_abund_ttest_sep, PLFA_abund_cohens_sep)
PLFA_abund_stats <- left_join(PLFA_abund_ttest, PLFA_abund_cohens)

#reformat output for joining for intensity data 
PLFA_abund_stats_rf_sep <- left_join(PLFA_df_abund_filt, PLFA_abund_stats_sep %>% select(c(Fraction, site, group1, group2, p, effsize)) %>% dplyr::rename("p_abund" = "p") %>% dplyr::rename("effsize_abund" = "effsize"))
PLFA_abund_stats_rf <- left_join(PLFA_df_abund_filt, PLFA_abund_stats %>% select(c(Fraction, group1, group2, p, effsize)) %>% dplyr::rename("p_abund" = "p") %>% dplyr::rename("effsize_abund" = "effsize"))
```

Quantification of the effect size magnitude is performed using the thresholds defined in Cohen (1992). The magnitude is assessed using the thresholds provided in (Cohen 1992), i.e. |d| < 0.2 "negligible", |d| < 0.5 "small", |d| < 0.8 "medium", otherwise "large".

## Combine 
```{r}
PLFA_stats_combined_sep <- left_join(PLFA_enrich_stats_rf_sep, PLFA_abund_stats_rf_sep)
PLFA_stats_combined <- left_join(PLFA_enrich_stats_rf, PLFA_abund_stats_rf)

#add categorical groupings 
PLFAs_classified <- PLFA_stats_combined %>% mutate(
  classification = case_when(
    #persistent: p_enrich > 0.05, abs(effsize_enrich) < 0.2, p_abund > 0.05, abs(effsize_abund) < 0.2
    p_enrich > 0.01 & abs(effsize_enrich) < 0.2 & p_abund > 0.01 & abs(effsize_abund) < 0.2 ~ "persistent",
    #continued production (labeled): p_abund < 0.05, effsize_abund > 0.5, p_enrich < 0.05, effsize_enrich > 0.5; effsize_abund > 0.5, p_enrich < 0.05, effsize_enrich < -0.5 (unlabeled)
    p_abund < 0.01 & effsize_abund > 0.8  ~ "continued production", 
    #degradation:  p_abund < 0.05, effsize_abund < -0.5, (old lipid) p_enrich < 0.05, effsize_enrich > 0.5, (new lipid) p_enrich < 0.05, effsize_enrich < -0.5
    p_abund < 0.01 & effsize_abund < -0.8  ~ "degraded", 
    TRUE ~ "unclassified"
  )
)
#separated 
PLFAs_classified_sep <- PLFA_stats_combined_sep %>% mutate(
    classification = case_when(
    #persistent: p_enrich > 0.05, abs(effsize_enrich) < 0.2, p_abund > 0.05, abs(effsize_abund) < 0.2
    p_enrich > 0.01 & abs(effsize_enrich) < 0.2 & p_abund > 0.01 & abs(effsize_abund) < 0.2 ~ "persistent",
    #continued production (labeled): p_abund < 0.05, effsize_abund > 0.5, p_enrich < 0.05, effsize_enrich > 0.5; effsize_abund > 0.5, p_enrich < 0.05, effsize_enrich < -0.5 (unlabeled)
    p_abund < 0.01 & effsize_abund > 0.8  ~ "continued production", 
    #degradation:  p_abund < 0.05, effsize_abund < -0.5, (old lipid) p_enrich < 0.05, effsize_enrich > 0.5, (new lipid) p_enrich < 0.05, effsize_enrich < -0.5
    p_abund < 0.01 & effsize_abund < -0.8  ~ "degraded", 
    TRUE ~ "unclassified"
  )
)
```

# Enrichment differences with site
```{r}
#perform t-test 
PLFA_enrichment_ttest_site <- PLFA_df_enrich_filt %>% group_by(Fraction) %>% mutate(n2 = sum(!is.na(wt.perc.13C.excess))) %>% filter(n2 == 20) %>% ungroup() %>%
  select(c(Fraction, timepoint, site, id, wt.perc.13C.excess)) %>%
  arrange(Fraction, timepoint, site, id) %>%
  mutate(
    site = factor(site), 
    timepoint = factor(timepoint)
  ) %>%
  group_by(Fraction) %>%
  t_test(wt.perc.13C.excess ~ site, paired = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
    
#calculate Cohen's d 
PLFA_enrichment_cohens_site <-PLFA_df_enrich_filt %>% group_by(Fraction) %>% mutate(n2 = sum(!is.na(wt.perc.13C.excess))) %>% filter(n2 == 20) %>% ungroup() %>%
  select(c(Fraction, timepoint, site, id, wt.perc.13C.excess)) %>%
  arrange(Fraction, timepoint, site, id) %>%
  mutate(
    site = factor(site), 
    timepoint = factor(timepoint)
  ) %>%
  group_by(Fraction) %>%
  cohens_d(wt.perc.13C.excess ~ site, paired = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))

#combine t-test and cohen's d outputs
PLFA_enrichment_stats_site <- left_join(PLFA_enrichment_ttest_site, PLFA_enrichment_cohens_site)

#reformat output for joining for intensity data 
PLFA_enrichment_stats_rf_site <- PLFA_enrichment_stats_site %>% select(c(Fraction, group1, group2, p, effsize)) %>% dplyr::rename("p_enrich" = "p") %>% dplyr::rename("effsize_enrich" = "effsize")
```


# Abundance differences with Site 
```{r}
#perform t-test 
PLFA_abund_ttest_site <- PLFA_df_abund_filt %>% group_by(Fraction) %>% mutate(n2 = sum(!is.na(ug.Total.C.per.g.dry.soil))) %>% filter(n2 == 20) %>% ungroup() %>%
  select(c(Fraction, timepoint, site, id, ug.Total.C.per.g.dry.soil)) %>%
  arrange(Fraction, timepoint, site, id) %>%
  mutate(
    site = factor(site), 
    timepoint = factor(timepoint)
  ) %>%
  group_by(Fraction) %>%
  t_test(ug.Total.C.per.g.dry.soil ~ site, paired = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))

#calculate Cohen's d 
PLFA_abund_cohens_site <- PLFA_df_abund_filt %>% group_by(Fraction) %>% mutate(n2 = sum(!is.na(ug.Total.C.per.g.dry.soil))) %>% filter(n2 == 20) %>% ungroup() %>%
  select(c(Fraction, timepoint, site, id, ug.Total.C.per.g.dry.soil)) %>%
  arrange(Fraction, timepoint, site, id) %>%
  mutate(
    site = factor(site), 
    timepoint = factor(timepoint)
  ) %>%
  group_by(Fraction) %>%
  cohens_d(ug.Total.C.per.g.dry.soil ~ site, paired = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))

#combine t-test and cohen's d outputs
PLFA_abund_stats_site <- left_join(PLFA_abund_ttest_site, PLFA_abund_cohens_site)

#reformat output for joining for intensity data 
PLFA_abund_stats_rf_site <- PLFA_abund_stats_site %>% select(c(Fraction, group1, group2, p, effsize)) %>% dplyr::rename("p_abund" = "p") %>% dplyr::rename("effsize_abund" = "effsize")
```

## Combine 
```{r}
PLFA_stats_combined_site <- left_join(PLFA_enrichment_stats_rf_site, PLFA_abund_stats_rf_site)

#add categorical groupings 
PLFAs_classified_site <- PLFA_stats_combined_site %>% mutate(
  site_classification = case_when(
    #both enrichment and intensity -> need to break down into sites further 
    p_enrich < 0.05 & effsize_enrich > 0.2 & p_abund < 0.05 & effsize_abund > 0.2 ~ "Wisconsin enrichment and intensity",
    p_enrich < 0.05 & effsize_enrich < -0.2 & p_abund < 0.05 & effsize_abund < -0.2 ~ "Michigan enrichment and intensity",
    p_enrich < 0.05 & effsize_enrich > 0.2 & p_abund < 0.05 & effsize_abund< -0.2 ~ "Wisconsin enrichment and Michigan intensity",
    p_enrich < 0.05 & effsize_enrich < -0.2 & p_abund < 0.05 & effsize_abund > 0.2 ~ "Michigan enrichment and Wisconsin intensity",
    #enrichment 
    p_enrich < 0.05 & effsize_enrich > 0.2  ~ "enrichment Wisconsin",
    p_enrich < 0.05 & effsize_enrich < -0.2  ~ "enrichment Michigan",
    #intensity 
    p_abund < 0.05 & effsize_abund > 0.2  ~ "intensity Wisconsin",
    p_abund < 0.05 & effsize_abund < 0.2  ~ "intensity Michigan",
      TRUE ~ "unclassified"
  )
  )

#add site classification to timepoint dataframe
#unseparated
all_classified <- left_join(PLFAs_classified, PLFAs_classified_site %>% select(c(Fraction, site_classification)))
#separated
all_classified_sep <- left_join(PLFAs_classified_sep, PLFAs_classified_site %>% select(c(Fraction, site_classification)))

#summarized 
#not separated by site
all_classified_summarized <- all_classified %>% 
mutate(log2_totalC = log2(ug.Total.C.per.g.dry.soil)) %>%
  filter(log2_totalC != -Inf) %>%
   group_by(Fraction) %>%
  mutate(
    excess_13C_avg = mean(wt.perc.13C.excess, na.rm = TRUE), 
    total_C_avg = mean(ug.Total.C.per.g.dry.soil, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  #need to pivot data to calculate ratio of intensities between timepoints 
  group_by(Fraction, timepoint) %>%
  mutate(
    excess_13C_timepoint_avg = mean(wt.perc.13C.excess, na.rm = TRUE),
    total_C_timepoint_avg = mean(ug.Total.C.per.g.dry.soil, na.rm = TRUE),
    mean_log2_C_timepoint = mean(log2_totalC, na.rm = TRUE)
  ) %>%
  ungroup() %>%
    pivot_wider(names_from = timepoint, values_from = c(total_C_timepoint_avg, excess_13C_timepoint_avg, mean_log2_C_timepoint)) %>%
  arrange(Fraction, id) %>%
  fill(total_C_timepoint_avg_1year, .direction = "up") %>%
  fill(excess_13C_timepoint_avg_1year, .direction = "up") %>%
  fill(mean_log2_C_timepoint_1year, .direction = "up") %>%
  fill(total_C_timepoint_avg_2month, .direction = "down") %>%
  fill(excess_13C_timepoint_avg_2month, .direction = "down") %>%
  fill(mean_log2_C_timepoint_2month, .direction = "down") %>%
mutate(log2FC_total_C = mean_log2_C_timepoint_1year - mean_log2_C_timepoint_2month) %>%
  select(c(Fraction, excess_13C_avg, total_C_avg, log2FC_total_C, excess_13C_timepoint_avg_1year, p_enrich, effsize_enrich, p_abund, effsize_abund, classification, site_classification)) %>% distinct()

#separated by site
all_classified_sep_summarized <- all_classified_sep %>% 
  mutate(log2_totalC = log2(ug.Total.C.per.g.dry.soil)) %>%
  filter(log2_totalC != -Inf) %>%
   group_by(Fraction, site) %>%
  mutate(
    excess_13C_avg = mean(wt.perc.13C.excess, na.rm = TRUE), 
    total_C_avg = mean(ug.Total.C.per.g.dry.soil, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  #need to pivot data to calculate ratio of abundances between timepoints 
  group_by(Fraction, site, timepoint) %>%
  mutate(
    excess_13C_timepoint_avg = mean(wt.perc.13C.excess, na.rm = TRUE),
    total_C_timepoint_avg = mean(ug.Total.C.per.g.dry.soil, na.rm = TRUE),
    mean_log2_C_timepoint = mean(log2_totalC, na.rm = TRUE)
  ) %>%
  ungroup() %>%
    pivot_wider(names_from = timepoint, values_from = c(total_C_timepoint_avg, excess_13C_timepoint_avg, mean_log2_C_timepoint)) %>%
  arrange(Fraction, id) %>%
  fill(total_C_timepoint_avg_1year, .direction = "up") %>%
  fill(excess_13C_timepoint_avg_1year, .direction = "up") %>%
  fill(mean_log2_C_timepoint_1year, .direction = "up") %>%
  fill(total_C_timepoint_avg_2month, .direction = "down") %>%
  fill(excess_13C_timepoint_avg_2month, .direction = "down") %>%
  fill(mean_log2_C_timepoint_2month, .direction = "down") %>%
  mutate(log2FC_total_C = mean_log2_C_timepoint_1year - mean_log2_C_timepoint_2month) %>%
  select(c(Fraction, site, excess_13C_avg, total_C_avg, log2FC_total_C, excess_13C_timepoint_avg_1year, p_enrich, effsize_enrich, p_abund, effsize_abund, classification, site_classification)) %>% distinct()
```

# Create Heatmap Tables 
## Add double bond and chain info
```{r}
plfa_meta <- readRDS("output/plfa_meta.rds")

all_summarized_w_meta <- left_join(all_classified_summarized, plfa_meta)
all_summarized_sep_w_meta <- left_join(all_classified_sep_summarized, plfa_meta)
all_w_meta <- left_join(all_classified, plfa_meta)
```

## Prepare dataframe for Complexheatmap
```{r}
#reorder data (MI switchgrass 2 month, MI switchgrass 1 year...then WI with same pattern)
PLFAs_wmeta_ordered <- all_summarized_w_meta %>% select(c(Fraction, Type, chain_C, chain_DBE, classification, excess_13C_avg, total_C_avg, log2FC_total_C, excess_13C_timepoint_avg_1year, p_enrich, effsize_enrich, p_abund, effsize_abund, classification, site_classification)) %>% 
#remove any rows with NA values 
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("persistent", "continued production", "degraded",  "unclassified"))) %>%
  group_by(classification, Type) %>% arrange(classification, effsize_enrich) %>% distinct()
#%>% mutate(atomperc_avg = atomperc_avg * 100, atomperc_avg_1year = atomperc_timepoint_avg_1year * 100)
#sep
#reorder data (MI switchgrass 2 month, MI switchgrass 1 year...then WI with same pattern)
PLFAs_wmeta_ordered_sep <- all_summarized_sep_w_meta %>% select(c(Fraction, Type, chain_C, chain_DBE, classification, excess_13C_avg, total_C_avg, log2FC_total_C, excess_13C_timepoint_avg_1year, p_enrich, effsize_enrich, p_abund, effsize_abund, classification, site, site_classification)) %>% 
  dplyr::rename("site_desig" = "site") %>%
#remove any rows with NA values 
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("persistent", "continued production", "degraded",  "unclassified"))) %>%
  group_by(classification, Type, site_desig) %>% arrange(classification, effsize_enrich) 

#remove "unclassified"
just_classified <- PLFAs_wmeta_ordered %>% filter(classification != "unclassified") 
just_classified_sep <- PLFAs_wmeta_ordered_sep %>% filter(classification != "unclassified") 

#row grouping vector
classification = just_classified$classification
classification_sep = just_classified_sep$classification
```

## Tables
### Set colors
```{r}
col_enrich_fun = circlize::colorRamp2(c(0, 12), c("white", "#FF7F00"))

col_C_fun = circlize::colorRamp2(c(12, 20), c("#FB9A99", "#E31A1C"))

col_DBE_fun = circlize::colorRamp2(c(0, 4), c("white", "#1F78B4"))

RColorBrewer::brewer.pal(9, "Set1")
Set1 <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF","#999999")
```

### All classifications, no site specific stats
```{r}
#enrichment
#matrix
mat_enrich <- as.matrix(just_classified$excess_13C_timepoint_avg_1year)

h_enrich <- Heatmap(mat_enrich, name = "excess wt% 13C", row_split = just_classified$classification, col = col_enrich_fun, row_title = NULL, column_title = expression("excess wt. %"^13*"C"[year]), cluster_rows = FALSE, cluster_row_slices = FALSE, column_title_side = "bottom")

#p value enrichment
p_enrich = as.vector(just_classified$p_enrich)
# label significant with asterisk
is_sig = p_enrich < 0.01
pch = rep("*", length(is_sig))
pch[!is_sig] = NA
# color mapping for -log10(pvalue)
pvalue_col_fun = circlize::colorRamp2(c(0, 2, 3), c("#33A02C", "white", "#6A3D9A")) 
ha2 = rowAnnotation(p = anno_simple(-log10(p_enrich), 
        col = pvalue_col_fun, pch = pch), annotation_name_rot = 0)
# generate two legends
# legend for pvalue
lgd_pvalue = Legend(title = "p-value", col_fun = pvalue_col_fun, at = c(0, 1, 2, 3), 
    labels = c("1", "0.1", "0.01", "0.001"))
# and one for the significant p-values
lgd_sig = Legend(pch = "*", type = "points", labels = "< 0.01")


#D value enrichment
effsize_enrich = as.vector(just_classified$effsize_enrich)
effsize_enrich <- round(effsize_enrich, digits = 1)
ha1 = rowAnnotation("d" = anno_numeric(effsize_enrich, 
        bg_gp = gpar(fill = ifelse(effsize_enrich > 0, "#33A02C", "#6A3D9A"), col = ifelse(effsize_enrich > 0, "#33A02C", "#6A3D9A"))), annotation_name_rot = 0)

#abundance barplot
ratio = as.vector(just_classified$log2FC_total_C)
ha3 = rowAnnotation("Log2FC C" = anno_barplot(ratio, baseline = 1, gp = gpar(fill = "#1F78B4", col = "#1F78B4"), width = unit(3, "cm")))

#p value abundance
p_abund = as.vector(just_classified$p_abund)
# label significant with asterisk
is_sig2 = p_abund < 0.01
pch2 = rep("*", length(is_sig2))
pch2[!is_sig2] = NA

ha5 = rowAnnotation(p = anno_simple(-log10(p_abund), 
        col = pvalue_col_fun, pch = pch2), annotation_name_rot = 0)

#D value abundance
effsize_abund = as.vector(just_classified$effsize_abund)
effsize_abund <- round(effsize_abund, digits = 1)
ha4 = rowAnnotation("d" = anno_numeric(effsize_abund, 
        bg_gp = gpar(fill = ifelse(effsize_abund > 0, "#33A02C", "#6A3D9A"), col = ifelse(effsize_abund > 0, "#33A02C", "#6A3D9A"))), annotation_name_rot = 0)


#Lipid type annotations 
#vector for colors 
class_colors <- just_classified %>% mutate(class_colors = case_when(
  Type == "Straight chained saturated" ~ "#E41A1C",
  Type == "Polyunsaturated" ~ "#FFFF33", 
),
class_colors = as.character(class_colors)) %>% ungroup() %>% select(c(class_colors, Type)) 
class_colors_vec = class_colors$class_colors

class_colors_uniq = class_colors %>% distinct() %>% mutate(Type = factor(Type, levels = c("Straight chained saturated", "Polyunsaturated"))) %>% arrange(Type)

ha_lipid = rowAnnotation(
  #class annotation 
    Class = anno_text(just_classified$`Fraction`, location = 0.5, just = "center", gp = gpar(fill = class_colors_vec, col = "black")))
lgd_lipid = Legend(title = "PLFA class", legend_gp = gpar(fill = class_colors_uniq$class_colors),
    labels = class_colors_uniq$Type)

 #list(bar = c("a" = "red", "b" = "green", "c" = "blue"))
  
#Lipid formula annotations 
ha_formula = rowAnnotation(
   #DBE
    `chain DBE` = just_classified$chain_DBE,
  #total C annotation 
    `chain C` = just_classified$chain_C,
    col = list(`chain DBE` = col_DBE_fun, `chain C` = col_C_fun)
)

#group by classification 
ha_class <- rowAnnotation(Classification = classification, show_annotation_name = FALSE, show_legend = FALSE,
    col = list(Classification = c("persistent" = "#CAB2D6", "continued production" = "#B2DF8A", "degraded" = "#A6CEE3")))
#label classifications
text_list = list(
    text1 = "degraded"
)
ha_empty = rowAnnotation(foo = anno_empty(border = FALSE, 
    width = unit(5, "mm")))
        #,annotation_legend_param = list(
            #classification = list(direction = "horizontal", title_position = "lefttop", nrow = 1)))


draw(ha_empty + ha_class + ha_lipid + ha_formula + h_enrich + ha1 + ha2 + ha3 + ha4 + ha5, annotation_legend_list = list(lgd_lipid, lgd_pvalue, lgd_sig), heatmap_legend_side = "right", annotation_legend_side = "right", merge_legends = TRUE)


for(i in 1:1) {
    decorate_annotation("foo", slice = i, {
        grid.text(paste(text_list[[i]], collapse = "\n"), x = unit(2, "mm"), rot = 90, just = "center")
    })
}
```

### All classifications, site specific stats
```{r}
#enrichment
#matrix
mat_enrich_sep <- as.matrix(just_classified_sep$excess_13C_timepoint_avg_1year)

h_enrich_sep <- Heatmap(mat_enrich_sep, name = "excess wt% 13C", row_split = just_classified_sep$classification, col = col_enrich_fun, row_title = NULL, column_title = expression("excess wt. %"^13*"C"[year]), cluster_rows = FALSE, cluster_row_slices = FALSE, column_title_side = "bottom", width = unit(1, "cm"))

#p value enrichment
p_enrich_sep = as.vector(just_classified_sep$p_enrich)
# label significant with asterisk
is_sig_sep = p_enrich_sep < 0.01
pch_sep = rep("*", length(is_sig_sep))
pch_sep[!is_sig_sep] = NA
# color mapping for -log10(pvalue)
pvalue_col_fun = circlize::colorRamp2(c(0, 2, 3), c("#33A02C", "white", "#6A3D9A")) 
ha2_sep = rowAnnotation(p = anno_simple(-log10(p_enrich_sep), 
        col = pvalue_col_fun, pch = pch_sep), annotation_name_rot = 0)
# generate two legends
# legend for pvalue
lgd_pvalue = Legend(title = "p-value", col_fun = pvalue_col_fun, at = c(0, 1, 2, 3), 
    labels = c("1", "0.1", "0.01", "0.001"))
# and one for the significant p-values
lgd_sig = Legend(pch = "*", type = "points", labels = "< 0.01")


#D value enrichment
effsize_enrich_sep = as.vector(just_classified_sep$effsize_enrich)
effsize_enrich_sep <- round(effsize_enrich_sep, digits = 1)
ha1_sep = rowAnnotation("d" = anno_numeric(effsize_enrich_sep, 
        bg_gp = gpar(fill = ifelse(effsize_enrich_sep > 0, "#33A02C", "#6A3D9A"), col = ifelse(effsize_enrich_sep > 0, "#33A02C", "#6A3D9A"))), annotation_name_rot = 0)

#abundance barplot
ratio_sep = as.vector(just_classified_sep$log2FC_total_C)
ha3_sep = rowAnnotation("Log2FC C" = anno_barplot(ratio_sep, baseline = 1, gp = gpar(fill = "#1F78B4", col = "#1F78B4"), width = unit(3, "cm")))

#p value abundance
p_abund_sep = as.vector(just_classified_sep$p_abund)
# label significant with asterisk
is_sig2_sep = p_abund_sep < 0.01
pch2_sep = rep("*", length(is_sig2_sep))
pch2_sep[!is_sig2_sep] = NA

ha5_sep = rowAnnotation(p = anno_simple(-log10(p_abund_sep), 
        col = pvalue_col_fun, pch = pch2_sep), annotation_name_rot = 0)

#D value abundance
effsize_abund_sep = as.vector(just_classified_sep$effsize_abund)
effsize_abund_sep <- round(effsize_abund_sep, digits = 1)
ha4_sep = rowAnnotation("d" = anno_numeric(effsize_abund_sep, 
        bg_gp = gpar(fill = ifelse(effsize_abund_sep > 0, "#33A02C", "#6A3D9A"), col = ifelse(effsize_abund_sep > 0, "#33A02C", "#6A3D9A"))), annotation_name_rot = 0)


#Lipid type annotations 
#vector for colors 
class_colors <- just_classified_sep %>% mutate(class_colors = case_when(
  Type == "Straight chained saturated" ~ "#E41A1C",
  #Type == "Polyunsaturated" ~ "#FFFF33", 
),
class_colors = as.character(class_colors)) %>% ungroup() %>% select(c(class_colors, Type)) 
class_colors_vec = class_colors$class_colors

class_colors_uniq = class_colors %>% distinct() %>% mutate(Type = factor(Type, levels = c("Straight chained saturated", "Polyunsaturated"))) %>% arrange(Type)

ha_lipid_sep = rowAnnotation(
  #class annotation 
    Class = anno_text(just_classified_sep$`Fraction`, location = 0.5, just = "center", gp = gpar(fill = class_colors_vec, col = "black")))
lgd_lipid_sep = Legend(title = "PLFA class", legend_gp = gpar(fill = class_colors_uniq$class_colors),
    labels = class_colors_uniq$Type)

 #list(bar = c("a" = "red", "b" = "green", "c" = "blue"))
  
#Lipid formula annotations 
ha_formula_sep = rowAnnotation(
   #DBE
    `chain DBE` = just_classified_sep$chain_DBE,
  #total C annotation 
    `chain C` = just_classified_sep$chain_C,
    col = list(`chain DBE` = col_DBE_fun, `chain C` = col_C_fun)
)

#group by classification 
ha_class_sep <- rowAnnotation(Classification = classification_sep, show_annotation_name = FALSE, show_legend = FALSE,
    col = list(Classification = c("persistent" = "#CAB2D6", "continued production" = "#B2DF8A", "degraded" = "#A6CEE3")))
#label classifications
text_list = list(
    text1 = "degraded"
)
ha_empty = rowAnnotation(foo = anno_empty(border = FALSE, 
    width = unit(5, "mm")))
        #,annotation_legend_param = list(
            #classification = list(direction = "horizontal", title_position = "lefttop", nrow = 1)))


draw(ha_empty + ha_class_sep + ha_lipid_sep + ha_formula_sep + h_enrich_sep + ha1_sep + ha2_sep + ha3_sep + ha4_sep + ha5_sep, annotation_legend_list = list(lgd_lipid_sep, lgd_pvalue, lgd_sig), heatmap_legend_side = "right", annotation_legend_side = "right", merge_legends = TRUE)


for(i in 1:1) {
    decorate_annotation("foo", slice = i, {
        grid.text(paste(text_list[[i]], collapse = "\n"), x = unit(2, "mm"), rot = 90, just = "center")
    })
}
```

### All classifications, no site specific stats, including unclassified 
```{r}
#row grouping vector
classification = PLFAs_wmeta_ordered$classification

#enrichment
#matrix
mat_enrich <- as.matrix(PLFAs_wmeta_ordered$excess_13C_timepoint_avg_1year)

h_enrich <- Heatmap(mat_enrich, name = "excess wt% 13C", row_split = PLFAs_wmeta_ordered$classification, col = col_enrich_fun, row_title = NULL, column_title = expression("excess wt. %"^13*"C"[year]), cluster_rows = FALSE, cluster_row_slices = FALSE, column_title_side = "bottom", width = unit(4, "cm"))

#p value enrichment
p_enrich = as.vector(PLFAs_wmeta_ordered$p_enrich)
# label significant with asterisk
is_sig = p_enrich < 0.01
pch = rep("*", length(is_sig))
pch[!is_sig] = NA
# color mapping for -log10(pvalue)
pvalue_col_fun = circlize::colorRamp2(c(0, 2, 3), c("#33A02C", "white", "#6A3D9A")) 
ha2 = rowAnnotation(p = anno_simple(-log10(p_enrich), 
        col = pvalue_col_fun, pch = pch), annotation_name_rot = 0)
# generate two legends
# legend for pvalue
lgd_pvalue = Legend(title = "p-value", col_fun = pvalue_col_fun, at = c(0, 1, 2, 3), 
    labels = c("1", "0.1", "0.01", "0.001"))
# and one for the significant p-values
lgd_sig = Legend(pch = "*", type = "points", labels = "< 0.01")


#D value enrichment
effsize_enrich = as.vector(PLFAs_wmeta_ordered$effsize_enrich)
effsize_enrich <- round(effsize_enrich, digits = 1)
ha1 = rowAnnotation("d" = anno_numeric(effsize_enrich, rg = c(-4,1),
        bg_gp = gpar(fill = ifelse(effsize_enrich > 0, "#33A02C", "#6A3D9A"), col = ifelse(effsize_enrich > 0, "#33A02C", "#6A3D9A")), align_to = 0, width = unit(2, "cm"), bar_width = unit(1, "npc") - unit(0.25, "npc"), labels_offset = unit(6, "mm"), labels_gp = gpar(fontsize = 10)), annotation_name_rot = 0)

#abundance barplot
ratio = as.vector(PLFAs_wmeta_ordered$log2FC_total_C)
ha3 = rowAnnotation("Log2FC C" = anno_barplot(ratio, baseline = 0, gp = gpar(fill = "#1F78B4", col = "#1F78B4"), width = unit(2, "cm")))

#p value abundance
p_abund = as.vector(PLFAs_wmeta_ordered$p_abund)
# label significant with asterisk
is_sig2 = p_abund < 0.01
pch2 = rep("*", length(is_sig2))
pch2[!is_sig2] = NA

ha5 = rowAnnotation(p = anno_simple(-log10(p_abund), 
        col = pvalue_col_fun, pch = pch2), annotation_name_rot = 0)

#D value abundance
effsize_abund = as.vector(PLFAs_wmeta_ordered$effsize_abund)
effsize_abund <- round(effsize_abund, digits = 1)
ha4 = rowAnnotation("d" = anno_numeric(effsize_abund,rg = c(-4, 1),
        bg_gp = gpar(fill = ifelse(effsize_abund > 0, "#33A02C", "#6A3D9A"), col = ifelse(effsize_abund > 0, "#33A02C", "#6A3D9A")), align_to = 0, width = unit(3, "cm"), bar_width = unit(1, "npc") - unit(0.25, "npc"), labels_offset = unit(8, "mm"), labels_gp = gpar(fontsize = 10)), annotation_name_rot = 0)


#Lipid type annotations 
#vector for colors 
class_colors <- PLFAs_wmeta_ordered %>% mutate(class_colors = case_when(
  Type == "Straight chained saturated" ~ "#E41A1C",
  Type == "Saturated with mid-chain" ~ "#377EB8",
  Type == "Terminally branched saturated" ~ "#4DAF4A",
  Type == "Monounsaturated" ~ "#984EA3",
  Type == "Branched monounsaturated" ~ "#FF7F00",
  Type == "Polyunsaturated" ~ "#FFFF33", 
  Type == "Cyclopropyl" ~ "#A65628",
  Type == "Hydroxylated" ~ "#F781BF"
),
class_colors = as.character(class_colors)) %>% ungroup() %>% select(c(class_colors, Type)) 
class_colors_vec = class_colors$class_colors

class_colors_uniq = class_colors %>% distinct() %>% mutate(Type = factor(Type, levels = c("Straight chained saturated", "Saturated with mid-chain", "Terminally branched saturated", "Monounsaturated", "Branched monounsaturated", "Polyunsaturated", "Cyclopropyl", "Hydroxylated"))) %>% arrange(Type)

ha_lipid = rowAnnotation(
  #class annotation 
    Class = anno_text(PLFAs_wmeta_ordered$`Fraction`, location = 0.5, just = "center", width = unit(8, "cm"), gp = gpar(fill = class_colors_vec, col = "black")))
lgd_lipid = Legend(title = "PLFA class", legend_gp = gpar(fill = class_colors_uniq$class_colors),
    labels = class_colors_uniq$Type)

 #list(bar = c("a" = "red", "b" = "green", "c" = "blue"))
  
#Lipid formula annotations 
ha_formula = rowAnnotation(
   #DBE
    `chain DBE` = PLFAs_wmeta_ordered$chain_DBE,
  #total C annotation 
    `chain C` = PLFAs_wmeta_ordered$chain_C,
    col = list(`chain DBE` = col_DBE_fun, `chain C` = col_C_fun)
)

#group by classification 
ha_class <- rowAnnotation(Classification = classification, show_annotation_name = FALSE, show_legend = FALSE,
    col = list(Classification = c("persistent" = "#CAB2D6", "continued production" = "#B2DF8A", "degraded" = "#A6CEE3", "unclassified" = "#F781BF")))
#label classifications
text_list = list(
    text1 = "degraded", 
    text2 = "unclassified"
)
ha_empty = rowAnnotation(foo = anno_empty(border = FALSE, 
    width = unit(5, "mm")))
        #,annotation_legend_param = list(
            #classification = list(direction = "horizontal", title_position = "lefttop", nrow = 1)))


draw(ha_empty + ha_class + ha_lipid + ha_formula + h_enrich + ha1 + ha2 + ha3 + ha4 + ha5, annotation_legend_list = list(lgd_lipid, lgd_pvalue, lgd_sig), heatmap_legend_side = "right", annotation_legend_side = "right", merge_legends = TRUE)


for(i in 1:2) {
    decorate_annotation("foo", slice = i, {
        grid.text(paste(text_list[[i]], collapse = "\n"), x = unit(2, "mm"), rot = 90, just = "center")
    })
}


```
