---
title: "Persistent Lipids"
author: "Kaitlin Rempfert"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged 
    number_sections: yes 
    css: stylesheet.css 
    toc: yes 
    toc_float: true 
    toc_depth: 3 
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# load required packages
library(tidyverse)
library(ggplot2)
library(pmartR)
library(pander)
library(emmeans)
library(openxlsx)
library(ggpubr)
library(rstatix)
library(Rodin)
library(ComplexHeatmap)
library(CHNOSZ)
```

# Enrichment Data
## Import enrichment data and lipid metadata
```{r}
e_data_enrich_long <- readRDS("output/e_data_enrich_long_meta.RDS")
lipidomics_Rodin_meta <- readRDS("output/lipidomics_Rodin_meta.Rds")
```


## Group IPLs by Rodin metaadata
```{r}
#add additional grouping variables
lipidomics_Rodin_meta <- lipidomics_Rodin_meta %>% mutate(
  `Sub class` = case_when(
   `Sub class` != "" ~ `Sub class`,
   `Double Bonds` == 0 ~ "saturated",
   `Double Bonds` == 1 ~ "monounsaturated",
   `Double Bonds` > 1 ~ "polyunsaturated"
  ),
  subclass = paste0(`Sub class`, " ", `Main class`)
)

#merge with metadata
e_data_enrich_long_Rodin <- left_join(e_data_enrich_long, lipidomics_Rodin_meta) %>% na.omit()

#calculate mean atom% of lipid groups
e_data_enrich_long_Rodin_g <- e_data_enrich_long_Rodin %>% group_by(subclass, id, TimePoint) %>% 
  filter(`Main class` != "CoQ") %>% mutate(
  atomperc_subclass = mean(atomperc, na.rm = TRUE)
) %>% ungroup() %>%
  group_by(`Sub class`, id, TimePoint) %>%
  mutate(atomperc_sub_class= mean(atomperc, na.rm = TRUE)) %>%
  ungroup() %>% 
  group_by(`Main class`, id, TimePoint) %>% 
  mutate(atomperc_headgroup = mean(atomperc, na.rm = TRUE)) %>% 
  ungroup() %>%
  group_by(Category, id, TimePoint) %>% 
  mutate(atomperc_category = mean(atomperc, na.rm = TRUE)) %>%
  ungroup() 

#reformat to long df for t-tests
e_data_enrich_longer_Rodin_g <- e_data_enrich_long_Rodin_g %>% 
  pivot_longer(cols = c(contains("atomperc_")), names_to = "group", names_prefix = "atomperc_", values_to = "atomperc_group") %>% 
  #add "Lipid Group" column to print the group name for group atomperc values
  mutate(
    `Lipid Group` = case_when(
      group == "subclass" ~ subclass, 
      group == "sub_class" ~ `Sub class`,
      group == "headgroup" ~ `Main class`,
      group == "category" ~ Category
    )
  )

e_data_enrich_long_meta <- e_data_enrich_longer_Rodin_g %>% select(c(`Lipid Group`, Site, TimePoint, id, group, atomperc_group)) %>% dplyr::rename("Compound" = "Lipid Group", "atomperc" = "atomperc_group")
```


```{r}
#perform t-test (keep sites separate in case a lipid is persistent in one site and not another)
compound_enrichment_ttest_sep <- e_data_enrich_long_meta  %>%
  select(c(Compound, Site, TimePoint, id, atomperc)) %>%
  arrange(Compound, Site, TimePoint, id) %>%
  group_by(Compound, Site) %>%
  t_test(atomperc ~ TimePoint, paired = TRUE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
#perform t-test again without separating sites
compound_enrichment_ttest <- e_data_enrich_long_meta  %>%
  select(c(Compound, Site, TimePoint, id, atomperc)) %>%
  arrange(Compound, Site, TimePoint, id) %>%
  group_by(Compound) %>%
  t_test(atomperc ~ TimePoint, paired = TRUE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
    
#calculate Cohen's d (keep sites separate in case a lipid is persistent in one site and not another)
compound_enrichment_cohens_sep <- e_data_enrich_long_meta %>% 
  select(c(Compound, Site, TimePoint, id, atomperc)) %>%
  arrange(Compound, Site, TimePoint, id) %>%
  group_by(Compound, Site) %>%
  cohens_d(atomperc ~ TimePoint, paired = TRUE)  
#perform Cohen's d without separating sites 
compound_enrichment_cohens <- e_data_enrich_long_meta %>% 
  select(c(Compound, Site, TimePoint, id, atomperc)) %>%
  arrange(Compound, Site, TimePoint, id) %>%
  group_by(Compound) %>%
  cohens_d(atomperc ~ TimePoint, paired = TRUE)  

#combine t-test and cohen's d outputs
compound_enrichment_stats_sep <- left_join(compound_enrichment_ttest_sep, compound_enrichment_cohens_sep)
compound_enrichment_stats <- left_join(compound_enrichment_ttest, compound_enrichment_cohens)

#reformat output for joining for intensity data 
compound_enrichment_stats_rf_sep <- left_join(e_data_enrich_long_meta, compound_enrichment_stats_sep %>% select(c(Compound, Site, group1, group2, p, effsize)) %>% dplyr::rename("p_enrich" = "p") %>% dplyr::rename("effsize_enrich" = "effsize"))
compound_enrichment_stats_rf <- left_join(e_data_enrich_long_meta, compound_enrichment_stats %>% select(c(Compound, group1, group2, p, effsize)) %>% dplyr::rename("p_enrich" = "p") %>% dplyr::rename("effsize_enrich" = "effsize"))
```

# Intensity Data
## Import Intensity Data
```{r}
#read in abundance data 
e_data_long_intensity <- readRDS("output/e_data_long_meta_filt.RDS") %>% select(-group)
```


```{r}
#merge with metadata
e_data_long_Rodin <- left_join(e_data_long_intensity, lipidomics_Rodin_meta) %>% na.omit() %>% mutate(NOSC = 4 - ((4*C + H - 3*N - 2*O + 5*P)/C))

#calculate mean log2 intensity  of lipid groups
e_data_long_Rodin_g <- e_data_long_Rodin %>% group_by(subclass, id, TimePoint) %>%
  filter(`Main class` != "CoQ") %>% mutate(
  log2_intensity_subclass = mean(log2_intensity, na.rm = TRUE),
                                 NOSC_subclass = mean(NOSC, na.rm = TRUE)) %>% 
  ungroup() %>%
  group_by(`Sub class`, id, TimePoint) %>%
  mutate(log2_intensity_sub_class= mean(log2_intensity, na.rm = TRUE),
         NOSC_sub_class = mean(NOSC, na.rm = TRUE)) %>%
  ungroup() %>% 
  group_by(`Main class`, id, TimePoint) %>% 
  mutate(log2_intensity_headgroup = mean(log2_intensity, na.rm = TRUE),
         NOSC_headgroup = mean(NOSC, na.rm = TRUE)) %>% 
  ungroup() %>%
  group_by(Category, id, TimePoint) %>% 
  mutate(log2_intensity_category = mean(log2_intensity, na.rm = TRUE),
         NOSC_category = mean(NOSC, na.rm = TRUE)) %>%
  ungroup() 

#reformat to long df for t-tests
#log2_intensity
e_data_longer_Rodin_g <- e_data_long_Rodin_g %>% 
  pivot_longer(cols = c(contains("log2_intensity_")), names_to = "group", names_prefix = "log2_intensity_", values_to = "log2_intensity_group") %>% 
  #add "Lipid Group" column to print the group name for group log2_intensity values
  mutate(
    `Lipid Group` = case_when(
      group == "subclass" ~ subclass, 
      group == "sub_class" ~ `Sub class`,
      group == "headgroup" ~ `Main class`,
      group == "category" ~ Category
    )
  )
#NOSC
e_data_longer2_Rodin_g <- e_data_long_Rodin_g %>% 
  pivot_longer(cols = c(contains("NOSC_")), names_to = "group", names_prefix = "NOSC_", values_to = "NOSC_group") %>% 
  #add "Lipid Group" column to print the group name for group log2_intensity values
  mutate(
    `Lipid Group` = case_when(
      group == "subclass" ~ subclass, 
      group == "sub_class" ~ `Sub class`,
      group == "headgroup" ~ `Main class`,
      group == "category" ~ Category
    )
  )

e_data_longerer_Rodin_g <- left_join(e_data_longer_Rodin_g, e_data_longer2_Rodin_g) 

e_data_long_meta <- e_data_longerer_Rodin_g %>% select(c(`Lipid Group`, Site, TimePoint, id, group, log2_intensity_group, NOSC_group)) %>% dplyr::rename("Compound" = "Lipid Group", "log2_intensity" = "log2_intensity_group", "NOSC" = "NOSC_group") %>% distinct()
```


```{r}
#perform t-test (keep sites separate in case a lipid is persistent in one site and not another)
compound_intensity_ttest_sep <- e_data_long_meta  %>%
  select(c(Compound, Site, TimePoint, id, log2_intensity, NOSC)) %>%
  arrange(Compound, Site, TimePoint, id) %>%
  group_by(Compound, Site) %>%
  t_test(log2_intensity ~ TimePoint, paired = TRUE, var.equal = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
#perform t-test again without separating sites 
compound_intensity_ttest <- e_data_long_meta  %>%
  select(c(Compound, Site, TimePoint, id, log2_intensity, NOSC)) %>%
  arrange(Compound, Site, TimePoint, id) %>%
  group_by(Compound) %>%
  t_test(log2_intensity ~ TimePoint, paired = TRUE, var.equal = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
    
#calculate Cohen's d (keep sites separate in case a lipid is persistent in one site and not another)
compound_intensity_cohens_sep <-e_data_long_meta %>% 
  select(c(Compound, Site, TimePoint, id, log2_intensity)) %>%
  arrange(Compound, Site, TimePoint, id) %>%
  group_by(Compound, Site) %>%
 cohens_d(log2_intensity ~ TimePoint, paired = TRUE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
#recalculate without separating sites
compound_intensity_cohens <-e_data_long_meta %>% 
  select(c(Compound, Site, TimePoint, id, log2_intensity)) %>%
  arrange(Compound, Site, TimePoint, id) %>%
  group_by(Compound) %>%
 cohens_d(log2_intensity ~ TimePoint, paired = TRUE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))

#combine t-test and cohen's d outputs
compound_intensity_stats_sep <- left_join(compound_intensity_ttest_sep, compound_intensity_cohens_sep)
compound_intensity_stats <- left_join(compound_intensity_ttest, compound_intensity_cohens)

#reformat output for joining for intensity data 
compound_intensity_stats_rf_sep <- left_join(e_data_long_meta, compound_intensity_stats_sep %>% select(c(Compound, Site, group1, group2, p, effsize)) %>% dplyr::rename("p_intensity" = "p") %>% dplyr::rename("effsize_intensity" = "effsize"))
compound_intensity_stats_rf <- left_join(e_data_long_meta, compound_intensity_stats %>% select(c(Compound, group1, group2, p, effsize)) %>% dplyr::rename("p_intensity" = "p") %>% dplyr::rename("effsize_intensity" = "effsize"))
```

Quantification of the effect size magnitude is performed using the thresholds defined in Cohen (1992). The magnitude is assessed using the thresholds provided in (Cohen 1992), i.e. |d| < 0.2 "negligible", |d| < 0.5 "small", |d| < 0.8 "medium", otherwise "large".

## Combine 
```{r}
compound_stats_combined_sep <- left_join(compound_enrichment_stats_rf_sep, compound_intensity_stats_rf_sep) %>% distinct()
compound_stats_combined <- left_join(compound_enrichment_stats_rf, compound_intensity_stats_rf) %>% distinct()

#add categorical groupings 
compounds_classified <- compound_stats_combined %>% mutate(
  classification = case_when(
    #persistent: p_enrich > 0.05, abs(effsize_enrich) < 0.2, p_intensity > 0.05, abs(effsize_intensity) < 0.2
    p_enrich > 0.01 & abs(effsize_enrich) < 0.2 & p_intensity > 0.01 & abs(effsize_intensity) < 0.2 ~ "persistent",
    #continued production (labeled): p_intensity < 0.05, effsize_intensity > 0.5, p_enrich < 0.05, effsize_enrich > 0.5; effsize_intensity > 0.5, p_enrich < 0.05, effsize_enrich < -0.5 (unlabeled)
    p_intensity < 0.01 & effsize_intensity > 0.8  ~ "continued production", 
    #degradation:  p_intensity < 0.05, effsize_intensity < -0.5, (old lipid) p_enrich < 0.05, effsize_enrich > 0.5, (new lipid) p_enrich < 0.05, effsize_enrich < -0.5
    p_intensity < 0.01 & effsize_intensity < -0.8  ~ "degraded", 
    TRUE ~ "unclassified"
  )
) 
#separated 
compounds_classified_sep <- compound_stats_combined_sep %>% mutate(
    classification = case_when(
    #persistent: p_enrich > 0.05, abs(effsize_enrich) < 0.2, p_intensity > 0.05, abs(effsize_intensity) < 0.2
    p_enrich > 0.01 & abs(effsize_enrich) < 0.2 & p_intensity > 0.01 & abs(effsize_intensity) < 0.2 ~ "persistent",
    #continued production (labeled): p_intensity < 0.05, effsize_intensity > 0.5, p_enrich < 0.05, effsize_enrich > 0.5; effsize_intensity > 0.5, p_enrich < 0.05, effsize_enrich < -0.5 (unlabeled)
    p_intensity < 0.01 & effsize_intensity > 0.8  ~ "continued production", 
    #degradation:  p_intensity < 0.05, effsize_intensity < -0.5, (old lipid) p_enrich < 0.05, effsize_enrich > 0.5, (new lipid) p_enrich < 0.05, effsize_enrich < -0.5
    p_intensity < 0.01 & effsize_intensity < -0.8  ~ "degraded", 
    TRUE ~ "unclassified"
  )
) 
```

# Enrichment differences with site
```{r}
#perform t-test 
compound_enrichment_ttest_site <- e_data_enrich_long_meta %>% 
  select(c(Compound, TimePoint, Site, id, atomperc)) %>%
  arrange(Compound, TimePoint, Site, id) %>%
  group_by(Compound) %>%
  t_test(atomperc ~ Site, paired = FALSE, var.equal = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
    
#calculate Cohen's d 
compound_enrichment_cohens_site <- e_data_enrich_long_meta %>% 
  select(c(Compound, TimePoint, Site, id, atomperc)) %>%
  arrange(Compound, TimePoint, Site, id) %>%
  group_by(Compound) %>%
  cohens_d(atomperc ~ Site, paired = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))

#combine t-test and cohen's d outputs
compound_enrichment_stats_site <- left_join(compound_enrichment_ttest_site, compound_enrichment_cohens_site)

#reformat output for joining for intensity data 
compound_enrichment_stats_rf_site <- compound_enrichment_stats_site %>% select(c(Compound, group1, group2, p, effsize)) %>% dplyr::rename("p_enrich" = "p") %>% dplyr::rename("effsize_enrich" = "effsize")
```

# Intensity differences with Site 
```{r}
#perform t-test
compound_intensity_ttest_site <- e_data_long_meta  %>%
  select(c(Compound, Site, TimePoint, id, log2_intensity)) %>%
  arrange(Compound, TimePoint, Site, id) %>%
  group_by(Compound) %>%
  t_test(log2_intensity ~ Site, var.equal = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))
    
#calculate Cohen's d 
compound_intensity_cohens_site <- e_data_long_meta %>%
  select(c(Compound, Site, TimePoint, id, log2_intensity)) %>%
  arrange(Compound, Site, TimePoint, id) %>%
  group_by(Compound) %>%
 cohens_d(log2_intensity ~ Site, paired = FALSE) %>%
  add_significance(cutpoints = c(0, 1e-04, 0.001, 0.01, 0.1, 1),
  symbols = c("****", "***", "**", "*", "ns"))

#combine t-test and cohen's d outputs
compound_intensity_stats_site <- left_join(compound_intensity_ttest_site, compound_intensity_cohens_site)
#reformat output for joining for intensity data 
compound_intensity_stats_rf_site <- compound_intensity_stats_site %>% select(c(Compound, group1, group2, p, effsize)) %>% dplyr::rename("p_intensity" = "p") %>% dplyr::rename("effsize_intensity" = "effsize")
```


## Combine 
```{r}
compound_stats_combined_site <- left_join(compound_enrichment_stats_rf_site, compound_intensity_stats_rf_site) %>% distinct()

#add categorical groupings 
compounds_classified_site <- compound_stats_combined_site %>% mutate(
  site_classification = case_when(
    #both enrichment and intensity -> need to break down into sites further 
    p_enrich < 0.05 & effsize_enrich > 0.2 & p_intensity < 0.05 & effsize_intensity > 0.2 ~ "Wisconsin enrichment and intensity",
    p_enrich < 0.05 & effsize_enrich < -0.2 & p_intensity < 0.05 & effsize_intensity < -0.2 ~ "Michigan enrichment and intensity",
    p_enrich < 0.05 & effsize_enrich > 0.2 & p_intensity < 0.05 & effsize_intensity < -0.2 ~ "Wisconsin enrichment and Michigan intensity",
    p_enrich < 0.05 & effsize_enrich < -0.2 & p_intensity < 0.05 & effsize_intensity > 0.2 ~ "Michigan enrichment and Wisconsin intensity",
    #enrichment 
    p_enrich < 0.05 & effsize_enrich > 0.2  ~ "enrichment Wisconsin",
    p_enrich < 0.05 & effsize_enrich < -0.2  ~ "enrichment Michigan",
    #intensity 
    p_intensity < 0.05 & effsize_intensity > 0.2  ~ "intensity Wisconsin",
    p_intensity < 0.05 & effsize_intensity < 0.2  ~ "intensity Michigan",
      TRUE ~ "unclassified"
  )
  )

#add site classification to timepoint dataframe
#unseparated
all_classified <- left_join(compounds_classified, compounds_classified_site %>% select(c(Compound, site_classification)))
#separated
all_classified_sep <- left_join(compounds_classified_sep, compounds_classified_site %>% select(c(Compound, site_classification)))

#summarized
#not separated by site
all_classified_summarized <- all_classified %>% 
   group_by(Compound) %>%
  mutate(
    atomperc_avg = mean(atomperc, na.rm = TRUE), 
    log2_intensity_avg = mean(log2_intensity, na.rm = TRUE),
    NOSC_avg = mean(NOSC, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  #need to pivot data to calculate ratio of intensities between timepoints 
  group_by(Compound, TimePoint) %>%
  mutate(
    intensity_timepoint_avg = mean(log2_intensity, na.rm = TRUE),
    atomperc_timepoint_avg = mean(atomperc, na.rm = TRUE)
  ) %>%
  ungroup() %>%
    pivot_wider(names_from = TimePoint, values_from = c(intensity_timepoint_avg, atomperc_timepoint_avg)) %>%
  arrange(Compound, id) %>%
  fill(intensity_timepoint_avg_1year, .direction = "down") %>%
  fill(atomperc_timepoint_avg_1year, .direction = "down") %>%
  fill(intensity_timepoint_avg_2month, .direction = "up") %>%
  fill(atomperc_timepoint_avg_2month, .direction = "up") %>%
  mutate(log2FC_intensity = intensity_timepoint_avg_2month - intensity_timepoint_avg_1year) %>%
  select(c(Compound, group, NOSC_avg, atomperc_avg, log2_intensity_avg, log2FC_intensity, atomperc_timepoint_avg_1year, p_enrich, effsize_enrich, p_intensity, effsize_intensity, classification, site_classification)) %>% distinct()

#separated by site
all_classified_sep_summarized <- all_classified_sep %>% 
   group_by(group, Compound, Site) %>%
  mutate(
    atomperc_avg = mean(atomperc, na.rm = TRUE), 
    log2_intensity_avg = mean(log2_intensity, na.rm = TRUE),
    NOSC_avg = mean(NOSC, na.rm =TRUE)
  ) %>%
  ungroup() %>%
  #need to pivot data to calculate ratio of intensities between timepoints 
  group_by(group, Compound, Site, TimePoint) %>%
  mutate(
    intensity_timepoint_avg = mean(log2_intensity, na.rm = TRUE),
    atomperc_timepoint_avg = mean(atomperc, na.rm = TRUE)
  ) %>%
  ungroup() %>%
    pivot_wider(names_from = TimePoint, values_from = c(intensity_timepoint_avg, atomperc_timepoint_avg)) %>%
  arrange(Compound, id) %>%
  fill(intensity_timepoint_avg_1year, .direction = "down") %>%
  fill(atomperc_timepoint_avg_1year, .direction = "down") %>%
  fill(intensity_timepoint_avg_2month, .direction = "up") %>%
  fill(atomperc_timepoint_avg_2month, .direction = "up") %>%
  mutate(log2FC_intensity = intensity_timepoint_avg_2month - intensity_timepoint_avg_1year) %>%
  select(c(Compound, group, Site, NOSC_avg, atomperc_avg, log2_intensity_avg, log2FC_intensity, atomperc_timepoint_avg_1year, p_enrich, effsize_enrich, p_intensity, effsize_intensity, classification, site_classification)) %>% distinct()
```

# Create Heatmap Tables 

# ComplexHeatmap Tables
## Prepare dataframe for Complexheatmap
```{r}
#reorder data (MI switchgrass 2 month, MI switchgrass 1 year...then WI with same pattern)
lipids_wmeta_ordered <- all_classified_summarized %>% select(c(Compound, group, classification, NOSC_avg, atomperc_avg, atomperc_timepoint_avg_1year, p_enrich, effsize_enrich, log2_intensity_avg, p_intensity, effsize_intensity, log2FC_intensity, site_classification)) %>% 
#remove any rows with NA values 
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("persistent", "continued production", "degraded",  "unclassified")),
         #add charge at neutral pH
         z = case_when(
           grepl("PI", Compound) ~ -1,
           grepl("hydroxy", Compound) ~ -1,
           grepl("PG", Compound) ~ -1,
           TRUE ~ 0
         )
         ) %>%
  group_by(classification, group) %>% arrange(classification, effsize_enrich) 
#%>% mutate(atomperc_avg = atomperc_avg * 100, atomperc_avg_1year = atomperc_timepoint_avg_1year * 100)
#sep
#reorder data (MI switchgrass 2 month, MI switchgrass 1 year...then WI with same pattern)
lipids_wmeta_ordered_sep <- all_classified_sep_summarized %>% select(c(Compound, group, classification, NOSC_avg, Site, atomperc_avg, atomperc_timepoint_avg_1year, p_enrich, effsize_enrich, log2_intensity_avg, p_intensity, effsize_intensity, log2FC_intensity, site_classification)) %>% 
  dplyr::rename("site_desig" = "Site") %>%
#remove any rows with NA values 
  na.omit() %>%
  mutate(classification = factor(classification, levels = c("persistent", "continued production", "degraded",  "unclassified"))) %>%
  group_by(classification, group, site_desig) %>% arrange(classification, effsize_enrich) 

#remove "unclassified"
just_classified <- lipids_wmeta_ordered %>% filter(classification != "unclassified") 
just_classified_sep <- lipids_wmeta_ordered_sep %>% filter(classification != "unclassified") 

#row grouping vector
classification = just_classified$classification
classification_sep = just_classified_sep$classification
```


## Tables
### Set colors
```{r}
col_atm_fun = circlize::colorRamp2(c(0, 0.015), c("white", "#FF7F00"))
#col_atm <- col_atm_fun(seq(-0.5, 5))

col_C_fun = circlize::colorRamp2(c(15, 20), c("white", "#E31A1C"))

col_DBE_fun = circlize::colorRamp2(c(0, 4), c("white", "#1F78B4"))

col_NOSC_fun = circlize::colorRamp2(c(-1.9, -1.3), c("#E31A1C", "white"))

col_z_fun = circlize::colorRamp2(c(-1, 1), c("#1F78B4", "white"))
```


### All subclassifications, no site specific stats 
```{r}
#create df of just subclasses 
lipidsub_wmeta_ordered <- lipids_wmeta_ordered %>% filter(group == "subclass") %>% 
  #add back `Main class` column for color scheme
  separate(Compound, c(NA, "Main class"), remove = FALSE)

classification = lipidsub_wmeta_ordered$classification 

#atom percent
#matrix
mat_atmperc <- as.matrix(lipidsub_wmeta_ordered$atomperc_timepoint_avg_1year)

h_atmperc2 <- Heatmap(mat_atmperc, name = "EAF", row_split = lipidsub_wmeta_ordered$classification, col = col_atm_fun, row_title = NULL, column_title = expression(""^13*"C EAF"[year]), cluster_rows = FALSE, cluster_row_slices = FALSE, column_title_side = "bottom", width = unit(3, "cm"))

#p value enrichment
p_enrich = as.vector(lipidsub_wmeta_ordered$p_enrich)
# label significant with asterisk
is_sig = p_enrich < 0.01
pch = rep("*", length(is_sig))
pch[!is_sig] = NA
# color mapping for -log10(pvalue)
pvalue_col_fun = circlize::colorRamp2(c(0, 2, 3), c("#33A02C", "white", "#6A3D9A")) 
ha2 = rowAnnotation(p = anno_simple(-log10(p_enrich), 
        col = pvalue_col_fun, pch = pch), annotation_name_rot = 0)
# generate two legends
# legend for pvalue
lgd_pvalue = Legend(title = "p-value", col_fun = pvalue_col_fun, at = c(0, 1, 2, 3), 
    labels = c("1", "0.1", "0.01", "0.001"))
# and one for the significant p-values
lgd_sig = Legend(pch = "*", type = "points", labels = "< 0.01")


#D value enrichment
effsize_enrich = as.vector(lipidsub_wmeta_ordered$effsize_enrich)
effsize_enrich <- round(effsize_enrich, digits = 1)
ha1 = rowAnnotation("d" = anno_numeric(effsize_enrich, 
        bg_gp = gpar(fill = ifelse(effsize_enrich > 0, "#33A02C", "#6A3D9A"), col = ifelse(effsize_enrich > 0, "#33A02C", "#6A3D9A"))), annotation_name_rot = 0)

#intensity barplot
ratio = as.vector(lipidsub_wmeta_ordered$log2FC_intensity)
ha3 = rowAnnotation("Log2FC intensity" = anno_barplot(ratio, baseline = 0, gp = gpar(fill = "#1F78B4", col = "#1F78B4"), width = unit(3, "cm")))

#NOSC
haN = rowAnnotation(
  "NOSC" = lipidsub_wmeta_ordered$NOSC_avg,
  col = list(NOSC = col_NOSC_fun)
)

#z
haz = rowAnnotation(
  "Z" = lipidsub_wmeta_ordered$z,
  col = list(Z = col_z_fun))


#p value intensity
#p value enrichment
p_intensity = as.vector(lipidsub_wmeta_ordered$p_intensity)
# label significant with asterisk
is_sig2 = p_intensity < 0.01
pch2 = rep("*", length(is_sig2))
pch2[!is_sig2] = NA

ha5 = rowAnnotation(p = anno_simple(-log10(p_intensity), 
        col = pvalue_col_fun, pch = pch2), annotation_name_rot = 0)

#D value intensity 
effsize_intensity = as.vector(lipidsub_wmeta_ordered$effsize_intensity)
effsize_intensity <- round(effsize_intensity, digits = 1)
ha4 = rowAnnotation("d" = anno_numeric(effsize_intensity, 
        bg_gp = gpar(fill = ifelse(effsize_intensity > 0, "#33A02C", "#6A3D9A"), col = ifelse(effsize_intensity > 0, "#33A02C", "#6A3D9A"))), annotation_name_rot = 0)


#Lipid type annotations 
#vector for colors 
class_colors <- lipidsub_wmeta_ordered %>% mutate(class_colors = case_when(
  `Main class` == "DG" ~ "#FDBF6F",
  `Main class` == "FA" ~ "#CAB2D6", 
  `Main class` == "PC" ~ "#A6CEE3", 
  `Main class` == "PE" ~ "#1F78B4", 
  `Main class` == "PG" ~ "#B2DF8A", 
  `Main class` == "PI" ~ "#33A02C", 
  `Main class` == "TG" ~ "#FF7F00", 
  `Main class` == "Cer" ~"#FB9A99", 
  `Main class` == "DGTSA" ~ "#E31A1C",
),
class_colors = as.character(class_colors)) %>% ungroup() %>% select(c(class_colors, `Main class`)) 
class_colors_vec = class_colors$class_colors

class_colors_uniq = class_colors %>% distinct() %>% mutate(`Main class` = factor(`Main class`, levels = c("PC", "PE", "PG", "PI", "Cer", "DGTSA", "DG", "TG", "FA"))) %>% arrange(`Main class`)

ha_lipid = rowAnnotation(
  #class annotation 
    Class = anno_text(lipidsub_wmeta_ordered$`Compound`, location = 0.5, just = "center", gp = gpar(fill = class_colors_vec, col = "white")))
lgd_lipid = Legend(title = "Lipid class", legend_gp = gpar(fill = class_colors_uniq$class_colors),
    labels = class_colors_uniq$`Main class`)

#group by classification 
ha_class <- rowAnnotation(Classification = classification, show_annotation_name = FALSE, show_legend = FALSE,
    col = list(Classification = c("persistent" = "#CAB2D6", "continued production" = "#B2DF8A", "degraded" = "#A6CEE3", "unclassified" = "#F781BF")))
#label classifications
text_list = list(
    text1 = "persistent",
    text2 = "continued production",
    text3 = "degraded", 
    text4 = "unclassified"
)
ha_empty = rowAnnotation(foo = anno_empty(border = FALSE, 
    width = unit(5, "mm")))
        #,annotation_legend_param = list(
            #classification = list(direction = "horizontal", title_position = "lefttop", nrow = 1)))


draw(ha_empty + ha_class + ha_lipid + haz + haN + h_atmperc2 + ha1 + ha2 + ha3 + ha4 + ha5, annotation_legend_list = list(lgd_lipid, lgd_pvalue, lgd_sig), heatmap_legend_side = "right", annotation_legend_side = "right", merge_legends = TRUE)


for(i in 1:4) {
    decorate_annotation("foo", slice = i, {
        grid.text(paste(text_list[[i]], collapse = "\n"), x = unit(6.5, "mm"), rot = 0, just = "right")
    })
}
```


