---
title: "TarMet_SIP"
output: html_document
date: '2022-11-01'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#devtools::install_github("hcji/TarMet")
library(TarMet)
# load required packages
library(tidyverse)
library(ggplot2)
library(stringi)
library(xcms)
library(IsoCorrectoR)
library(rstatix)
library(ggpubr)
library(MSnbase)

#import scripts here
source(file.path("scripts", "XIC.R"))
```


## Import Data 
```{r}
#path to positive .mzXML files
dda_folder_pos = "~/Documents/EC/data/incubation/lipidome/mzXML/positive/batch3_samples"
#list all files inside specified folder
rawfiles_pos <- list.files(dda_folder_pos, recursive = TRUE, full.names = TRUE, pattern = "\\.mzXML$")

#load into TarMet
samples_pos <- sapply(strsplit(rawfiles_pos, '/'), function(s) s[length(s)])
rawDataset_pos <- lapply(rawfiles_pos, LoadData_KRR)

#import metadata file 
meta <- readxl::read_excel("incubation_metadata_w_qc.xlsx")
#filter metadata to just batch 3
meta_samp_pos <- meta %>% filter(batch == "3" & polarity == "positive") %>% filter(!is.na(site)) 
#add columns to metadata
meta_samp_pos <- meta_samp_pos %>% mutate(
  #create label of labeled/unlabeled and rep (block)
  label_group = ifelse(amendment == "13C_Gluc", "L", "U"),
  label_group = paste0(label_group, block)
)

#create dataframe with file extensions and sample names
pd_samp_pos <- data.frame(extension = rawfiles_pos, sample_name = basename(rawfiles_pos))
#merge metadata with phenodata frame 
pd_samp_pos <- left_join(meta_samp_pos, pd_samp_pos, by = "sample_name")
#export ordered file list
samps_pos <- as.vector(pd_samp_pos$extension)

#load into xcms
dda_data_pos <- readMSData(rawfiles_pos,  pdata = new("NAnnotatedDataFrame", pd_samp_pos), centroided = TRUE, mode = "onDisk")
#set peak-picking parameters
cwp <- CentWaveParam(A = 4.289723e-07, ppm=1, Instrument=2, peakwidth=c(2.5, 45), snthresh = 10, noise=10000, prefilter=c(3, 100000), firstBaselineCheck = FALSE, integrate=2)
#pick peaks
dda_data_peaks_pos <- findChromPeaks(dda_data_pos, param = cwp)
```

Custom script to load files
```{r}
LoadData_KRR <- function(filename)
{
  msobj <- openMSfile(filename,backend="pwiz")
  peakInfo <- peaks(msobj)
  headerInfo <- header(msobj)
  whMS1 <- which(headerInfo$msLevel==1)
  peakInfo <- peakInfo[whMS1]
  
  peakInfo <- lapply(peakInfo, function(spectrum) {
    keep <- spectrum[,2] > 1e-6
    output <- as.data.frame(spectrum[keep,,drop = FALSE])
    colnames(output) <- c('mz','intensity')
    return(output)
  })
  
  scanTime <- round(headerInfo$retentionTime[whMS1],3)
  # close(msobj)
  
  return(list(path=filename, times=scanTime, peaks = peakInfo))
}
```

#create config file
```{r}
#import database of lipids identified in LIQUID
batch3_LIQUID <- read.csv("output/LIQUID/pos_LIQUID_db.csv") %>% select(-X)

### FOR EXTRACTING PEAK WIDTHS
#pull out retention times 
rt_targets_pos <- batch3_LIQUID %>% select(-c(Formula, Category, Adduct)) %>% dplyr::rename(compound = Compound) %>% mutate(rt_window = 30) 
#extract peak intensity and retention time
peak_param_pos <- extract_peak_param(rt_targets_pos, dda_data_peaks_pos)
#convert intensity and retention time columns to class numeric for filtering
peak_param_pos <- peak_param_pos %>% ungroup() %>% mutate(Intensity = as.numeric(as.character(Intensity)), RT = as.numeric(as.character(RT)), RT_width = as.numeric(as.character(RT_width)), compound = as.character(compound))
#except for FA that vary significantly in RT, average RT; average RT_width for all
peak_param_avg_pos <- peak_param_pos %>% group_by(compound) %>% mutate(
  targetRtAxis = mean(RT, na.rm = TRUE) * 60,
  RT_width = mean(RT_width, na.rm = TRUE)) %>% 
  mutate(
    targetRtLeft = targetRtAxis - RT_width/2,
    targetRtRight = targetRtAxis + RT_width/2
  ) %>% select(-c(RT_width, RT)) %>% 
  #select more intense natural abundance peak to be used for quality checking
  filter(amendment == "NA_Gluc") %>% select(-amendment) %>% group_by(compound) %>%
  slice_max(Intensity) %>% select(-Intensity) #replace catalog_number where intensity is greatest with the index from the vector
SampInd_pos <- samples_pos %>% as.data.frame() %>% mutate(sampleInd = row_number(), catalog_number = str_extract(., "[^_]+")) %>% select(-.)
#rejoin
peak_info_pos <- left_join(peak_param_avg_pos, SampInd_pos) %>% select(-c(catalog_number))

#export chemical formulas from database
batch3_db <- batch3_LIQUID %>% select(c(Compound, Formula, Adduct, mz_db)) %>% 
  #string split the chemical formulas
  mutate(
    name=str_extract_all(Formula,"[A-Za-z]+"),
    value=str_extract_all(Formula,"\\d+")
    ) %>%
   unnest() %>% spread(name,value = value, fill=0) 
#add random string as Identifier 
batch3_db$Identifier <- stringi::stri_rand_strings(length(unique(batch3_db$Compound)), 5, pattern = "[a-z]")

#subtract adduct from chemical formula
batch3_db_adductremoved <- batch3_db %>% mutate(
  H = ifelse(Adduct == "[M+NH4]+", as.numeric(H) - 4, as.numeric(H) - 1), 
  N = ifelse(Adduct == "[M+NH4]+", as.numeric(N) - 1, N), 
  N = as.numeric(N), 
  P = as.numeric(P),
  Formula_no_adduct = case_when(
    N != 0 & P != 0 ~ paste0("C", C, "H", H, "N", N, "O", O, "P", P),
    N != 0 & P == 0 ~ paste0("C", C, "H", H, "N", N, "O", O),
    N == 0 & P == 0 ~ paste0("C", C, "H", H, "O", O), 
    TRUE ~ "formula error")
)

#create config file
config_pos <- left_join(batch3_db_adductremoved, peak_info_pos, by = c("Compound" = "compound")) %>% mutate(adduct = ifelse(Adduct == "[M+NH4]+", "M+NH4", "M+H")) %>% select(c(Compound, Identifier, mz_db, Formula_no_adduct, adduct, targetRtAxis, targetRtLeft, targetRtRight, sampleInd, C)) %>% dplyr::rename("name" = "Compound", "mz" = "mz_db", "formula" = "Formula_no_adduct") %>% mutate(
  rtmin = targetRtAxis - 60,
  rtmax = targetRtAxis + 60,
  ppm = 5, 
  scale = 5, 
  height = 1000, 
  snr = 100, 
  tracer_element = 'C',     # type of element
  tracer_isotope = '13C',     # type of isotope tracer
  tracer_number = round(as.numeric(C)*.5, 0)
) %>% select(-C) %>% filter(name != "PC(20:4/20:5)") %>% filter(targetRtAxis != "NaN")

saveRDS(config_pos, "output/TarMet/pos/config_pos_4_5_23.rds")
```

Function for running TarMet using a config file
```{r}
#function for running TarMet using a config file
run_TarMet_fromconfig <- function(config, rawfiles, samples, rawDataset, file_extension) {
  output_total = data.frame()
  for (i in 1:nrow(config)){
    # set global parameters
      fineness <- 'Medium'
      resolution <- 60000
      shift <- 20  # parameter for alignment
      segment <- 20 
      # set the parameters for the specific metabolite
      threshold <- 0.01
      name <- config$name[i]
      Identifier <- config$Identifier[i]
      formula <- config$formula[i]
      mz <- config$mz[i]
      sampleInd <- config$sampleInd[i]
      tracer_element <- config$tracer_element[i]
      tracer_isotope <- config$tracer_isotope[i]
      tracer_number <-  config$tracer_number[i]
      targetRtAxis <- config$targetRtAxis[i]
      targetRtLeft <- config$targetRtLeft[i]
      targetRtRight <- config$targetRtRight[i]
      adduct <- config$adduct[i]
      ppm <- config$ppm[i] 
      scale <- config$scale[i]
      rtmin <- config$rtmin[i]
      rtmax <- config$rtmax[i]
      height <- config$height[i]
      snr <- config$snr[i]
      ## Calculate m/z
       mzs <- getMzWithTracer(formula, adduct, tracer_element, tracer_isotope, tracer_number)
      targetMzRanges <- getMzRanges(mzs, resolution=60000, ppm=ppm)   
    ## Targeted EIC generation
      targetEICs <- list()
      rtranges <- c(rtmin, rtmax)
      for (j in seq_along(rawfiles)){
        targetEICs[[j]] <- getMzEICs(rawDataset[[j]], rtranges=rtranges,
                                     mzranges=targetMzRanges, baseline=FALSE,
                                     smooth=TRUE)
        }
    pattern <- getIsotopicPattern(formula, adduct, threshold = .01, resolution=60000)
    theoretical <- as.numeric(pattern[,2])
    targetPeaks <- getIsotopicPeaks(targetEICs[[sampleInd]], SNR.Th=snr, peakScaleRange=scale, peakThr=height, theoretical=theoretical, fineness=fineness)
    #plot EICs
    p1 <- plotEICs(targetEICs[[sampleInd]], targetPeaks)
    cmp_name <- config$name[i]
    file_name <- paste0(file_extension, "/EICs/", i, "_", config$Identifier[i], ".html")
htmlwidgets::saveWidget(widget = p1, file = file_name, title = cmp_name)
    #output peak info
    outputPeakInfo <- targetPeaks$PeakInfo
    ## Set peak bounds manually 
    userInfo <- data.frame(Name='User', Position=targetRtAxis, Start=targetRtLeft, End=targetRtRight)
    outputPeakInfo <- rbind(targetPeaks$PeakInfo[,1:ncol(userInfo)], userInfo)
    ## Calculate peak area
    userArea <- round(getArea(targetEICs[[sampleInd]], targetRtLeft, targetRtRight), 2)
    outputPeakArea <- {
      res <- targetPeaks$PeakArea
      res <- cbind(rownames(res), res)
      res <- cbind(res, userArea, round(userArea/sum(userArea),3))
      colnames(res)[1] <- 'mzRange'
      colnames(res)[(ncol(res)-1) : ncol(res)] <- c('User','Abundance')
      res
    }
    ## Alignment
    alignedEICs <- getAlignedEICs(targetEICs, sampleInd, shift, segment, align=TRUE)
    WhtoPlot <- 1  # plot the M+1 feature of the targeted metabolite of all samples
    EicstoPlot <- lapply(alignedEICs, function(eics){
      eics[[WhtoPlot]]
      })
    names(EicstoPlot) <- samples
    #plot alignments
    p2 <- plotEICs(EicstoPlot, rtrange=c(outputPeakInfo$Start[WhtoPlot], outputPeakInfo$End[WhtoPlot]))
file_name2 <- paste0(file_extension, "/alignment/", i, "_", config$Identifier[i], ".html")
htmlwidgets::saveWidget(widget = p2, file = file_name2, title = cmp_name)
## Compare Peak Areas between isotopic peaks
whPeak <- which(outputPeakInfo$Name == 'User')
outputSamplesArea <- getQuantifiedResult(alignedEICs, outputPeakInfo$Start[whPeak], outputPeakInfo$End[whPeak])
colnames(outputSamplesArea) <- samples
#plot isotope stack barplot
p3 <- plotStackBar(outputSamplesArea)
file_name3 <- paste0(file_extension, "/isotopestackbar/", i, "_", config$Identifier[i], ".html")
htmlwidgets::saveWidget(widget = p3, file = file_name3, title = cmp_name)
    measurements <- as.data.frame(outputSamplesArea) 
    rownames(measurements) <- NULL
    rownames(measurements) <- paste0(Identifier, "_", as.numeric(rownames(measurements)) - 1)
    output_total <- rbind(output_total, measurements)
  }
  return(output_total)
}
```
  
## run TarMet
```{r}
measurements_pos <- run_TarMet_fromconfig(config = pos_config, rawfiles = rawfiles_pos, samples = samples_pos, rawDataset = rawDataset_pos, file_extension = "output/TarMet/pos/SIP/4_5_23")

#manual inspect for max tracer_number (remove coeluting peaks) and update spreadsheet
tracer_number_pos_upd <- readxl::read_excel("output/TarMet/pos/SIP_pos_tracer_number_upd.xlsx") %>% select(-c(`...1`, `...2`, `...3`, tracer_number))

config_pos_upd <- cbind (pos_config, tracer_number_pos_upd) %>% select(-tracer_number) %>% dplyr::rename("tracer_number" = "tracer_number_upd") %>% filter(tracer_number > 2)

measurements_pos_upd <- run_TarMet_fromconfig(config = config_pos_upd, rawfiles = rawfiles_pos, samples = samples_pos, rawDataset = rawDataset_pos, file_extension = "output/TarMet/pos/SIP/4_5_23")
```


## remove data for samples where monoisotopic peak is not detected
```{r}
measurements_pos_upd_filt <- measurements_pos_upd %>% rownames_to_column("samples") %>% separate(samples, c("Identifier", "Channel")) %>% group_by(Identifier) %>% rowwise() %>% mutate_at(vars(starts_with("EC")), funs((case_when(Channel == 0 & . < 10000 ~ 0, TRUE ~ .)))) %>% mutate(samples = paste0(Identifier, "_", Channel)) %>% ungroup() %>% select(-c(Identifier, Channel)) %>% column_to_rownames("samples")
```

## prepare for IsoCorrectoR
```{r}
#measurement file
measurements_pos_file <- measurements_pos_upd_filt %>% rownames_to_column("Measurements/Samples")
write.csv(x = measurements_pos_file, file = "output/TarMet/pos/measurements_file_4_23.csv", quote = FALSE, row.names = FALSE, na = "")

#molecule file
#make molecules file (will need to subset)
molecules_pos <- batch3_db %>% select(c(Identifier, Formula, C)) %>% dplyr::rename(Molecule = Identifier) %>% mutate(Formula = paste0(Formula, "LabC", C)) %>% select(Molecule, Formula) %>% mutate(MS = "") %>%
  setNames(c("Molecule", "MS ion or MS/MS product ion", "MS/MS neutral loss")) 
write.csv(x = molecules_pos, file = "output/TarMet/pos/molecule_file_4_23.csv", quote = FALSE, row.names = FALSE, na = "")

#make key for molecules file
key_pos <- batch3_db %>% select(Identifier, Compound)
write.csv(x = key_pos, file = "output/TarMet/pos/key_pos_4_23.csv", quote = FALSE, row.names = FALSE, na = "")
```

## Process data through IsoCorrectoR
```{r}
IsoCorrection(MeasurementFile = "output/TarMet/pos/measurements_file_4_23.csv", 
              ElementFile = "output/TarMet/pos/element_file.csv", 
              MoleculeFile = "output/TarMet/pos/molecule_file_4_23.csv",
              CorrectTracerImpurity = TRUE, CorrectTracerElementCore = TRUE, 
              CalculateMeanEnrichment = TRUE, UltraHighRes = FALSE, 
              DirOut = 'output/IsoCorrectoR/pos', 
              FileOut = 'result', 
              FileOutFormat = 'csv', 
              ReturnResultsObject = TRUE, CorrectAlsoMonoisotopic = FALSE, 
              CalculationThreshold = 10^-10, 
              CalculationThreshold_UHR = 8, 
              verbose = TRUE, Testmode = FALSE)
```


### Calculate paired rep (labeled vs unlabeled) relative mean enrichment & implement paired t-test with Cohen's D 
```{r}
#read in data from IsoCorrectoR
isoR_enrich_pos <- read.csv("output/IsoCorrectoR/pos/2023-05-02_170737/IsoCorrectoR_result_MeanEnrichment.csv", header = TRUE, sep = ",", quote = "\"", na.strings = "NA")
#Manipulate data into long format
isoR_enrich_long_pos <- tidyr::pivot_longer(data = isoR_enrich_pos, cols = starts_with("EC"), names_to = "sample_name", values_to = "mean.enrichment") %>% dplyr::rename(Identifier = X)
#join with metadata
isoR_enrich_long_pos <- left_join(isoR_enrich_long_pos, meta_samp_pos) %>%
#join with key to get compound names
left_join(key_pos) %>%
#add column "Label" to indicate labeling treatment (unlabeled vs labeled) and "Rep" with the rep #
mutate(Label = ifelse(grepl("L", label_group), "labeled", "unlabeled"), Rep = readr::parse_number(as.character(block)), Label = as.factor(Label))
saveRDS(isoR_enrich_long_pos, "output/isoR_enrich_long_pos_4_23.RDS")

#need to filter out reps that don't have compound detected in labeled and unlabeled as well as compounds that only show up in one paired rep
isoR_enrich_long_pos_filt <- isoR_enrich_long_pos %>% group_by(group, Identifier, Compound, Rep) %>% mutate(n = n()) %>% filter(n == 2) %>% ungroup() %>% group_by(group, Identifier, Compound) %>% mutate(n2 = n()) %>% filter(n2 > 2) %>% ungroup() %>% select(c(group, Identifier, Compound, mean.enrichment, Label, Rep))

#paired t-test
t_test_pos <- isoR_enrich_long_pos_filt %>%
  group_by(group, Identifier, Compound) %>%
  t_test(mean.enrichment ~ Label, paired = TRUE, detailed = TRUE) %>%
  add_significance() 
  
#calculate effect size (Cohen's D)
cohens_d_pos <- isoR_enrich_long_pos_filt %>% 
  group_by(group, Identifier, Compound) %>%
  cohens_d(mean.enrichment ~ Label, paired = TRUE)

#calculate relative enrichment between unlabeled and labeled reps
rel_enrich_pos <- isoR_enrich_long_pos_filt %>%
  group_by(group, Identifier, Compound, Rep) %>% 
  pivot_wider(names_from = Label, values_from = mean.enrichment) %>% 
  summarise(labeled = dplyr::first(na.omit(labeled)),
      unlabeled = dplyr::first(na.omit(unlabeled))) %>%
  mutate(rel_enrichment = labeled - unlabeled) %>%
#calculate mean + sd enrichment difference between unlabeled and labeled rep pairs
  group_by(group, Identifier, Compound) %>% 
  mutate(mean_rel_enrichment = mean(rel_enrichment), 
         sd_rel_enrichment = sd(rel_enrichment), Rep =Rep)

#combine
pos_SIP <- left_join(t_test_pos, cohens_d_pos) %>% left_join(rel_enrich_pos)
saveRDS(pos_SIP, "output/pos_SIP_4_23.RDS" )
```



#NEGATIVE

## Import Data 
```{r}
#path to .mzXML files
dda_folder_neg = "/Users/remp858/Documents/EC/data/incubation/lipidome/mzXML/negative/batch4_samples"
#list all files inside specified folder
rawfiles_neg <- list.files(dda_folder_neg, recursive = TRUE, full.names = TRUE, pattern = "\\.mzXML$")

#load into TarMet
samples_neg <- sapply(strsplit(rawfiles_neg, '/'), function(s) s[length(s)])
rawDataset_neg <- lapply(rawfiles_neg, LoadData_KRR)

####DO NOT NEED THIS BELOW UNLESS EXTRACTING PEAK PARAMETERS
#load into xcms
#import metadata file 
meta <- readxl::read_excel("incubation_metadata_w_qc.xlsx")
#filter metadata to just batch 3
meta_samp_neg <- meta %>% filter(batch == "4" & polarity == "negative") %>% filter(!is.na(site)) 
#add columns to metadata
meta_samp_neg <- meta_samp_neg %>% mutate(
  #create label of labeled/unlabeled and rep (block)
  label_group = ifelse(amendment == "13C_Gluc", "L", "U"),
  label_group = paste0(label_group, block)
)
#create dataframe witg file extensions and sample names
pd_samp_neg <- data.frame(extension = rawfiles_neg, sample_name = basename(rawfiles_neg))
#merge metadata with phenodata frame 
pd_samp_neg <- left_join(meta_samp_neg, pd_samp_neg, by = "sample_name")
#export ordered file list
samps_neg <- as.vector(pd_samp_neg$extension)

#load data 
dda_data_neg <- readMSData(samps_neg,  pdata = new("NAnnotatedDataFrame", pd_samp_neg), centroided = TRUE, mode = "onDisk")

#set peak-picking parameters
cwp <- CentWaveParam(ppm=1, peakwidth=c(2.5, 45), snthresh = 10, noise=10000, prefilter=c(3, 100000), firstBaselineCheck = FALSE, integrate=2)
#pick peaks
dda_data_peaks_neg <- findChromPeaks(dda_data_neg, param = cwp)
```

#create config file
```{r}
#import database of lipids identified in LIQUID
batch4_LIQUID <- read.csv("output/LIQUID/neg_LIQUID_db.csv") 
#import FA database
FA <- readxl::read_xlsx("output/LIQUID/fatty_acids.xlsx") %>% mutate(Category = "Fatty Acyls [FA]", Adduct = "[M-H]-") %>% dplyr::rename("Compound" = "compound")
#bind together databases
batch4_LIQUID <- rbind(batch4_LIQUID, FA)

### FOR EXTRACTING PEAK WIDTHS
#pull out retention times 
rt_targets_neg <- batch4_LIQUID %>% select(-c(Formula, Category, Adduct)) %>% dplyr::rename(compound = Compound) %>% mutate(rt_window = ifelse(grepl("FA", compound), 80, 30))
#extract peak intensity and retention time
peak_param_neg <- extract_peak_param(rt_targets_neg, dda_data_peaks_neg)
#convert intensity and retention time columns to class numeric for filtering
peak_param_neg <- peak_param_neg %>% ungroup() %>% mutate(Intensity = as.numeric(as.character(Intensity)), RT = as.numeric(as.character(RT)), RT_width = as.numeric(as.character(RT_width)), compound = as.character(compound))
#except for FA that vary significantly in RT, average RT; average RT_width for all
peak_param_avg_neg <- peak_param_neg  %>% group_by(compound) %>% mutate(
  targetRtAxis = mean(RT, na.rm = TRUE) * 60,
  RT_width = mean(RT_width, na.rm = TRUE)) %>% 
  mutate(
    targetRtLeft = targetRtAxis - RT_width/2,
    targetRtRight = targetRtAxis + RT_width/2
  ) %>% #select more intense natural abundance peak to be used for quality checking
  filter(amendment == "NA_Gluc") %>% select(-amendment) %>% group_by(compound) %>%
  slice_max(Intensity) %>% select(-Intensity) #replace catalog_number where intensity is greatest with the index from the vector
SampInd_neg <- samples_neg %>% as.data.frame() %>% mutate(sampleInd = row_number(), catalog_number = str_extract(., "[^_]+")) %>% select(-.) #replace catalog_number where intensity is greatest with the index from the vector
peak_info_neg <- left_join(peak_param_avg_neg, SampInd_neg) %>% select(-catalog_number)

#export chemical formulas from database
batch4_db <- batch4_LIQUID %>% select(c(Compound, Formula, Adduct, mz_db)) %>%  #string split the chemical formulas
  mutate(
    name=str_extract_all(Formula,"[A-Za-z]+"),
    value=str_extract_all(Formula,"\\d+")
    ) %>% unnest() %>% spread(name,value=value, fill=0)
#add random string as Identifier 
batch4_db$Identifier <- stringi::stri_rand_strings(length(unique(batch4_db$Compound)), 5, pattern = "[a-z]")

#subtract adduct from chemical formula
batch4_db_adductremoved <- batch4_db %>% mutate(
  H = as.numeric(H) + 1, 
  Formula_no_adduct = case_when(
    N != 0 & P != 0 ~ paste0("C", C, "H", H, "N", N, "O", O, "P", P),
    N != 0 & P == 0 ~ paste0("C", C, "H", H, "N", N, "O", O),
    N == 0 & P == 0 ~ paste0("C", C, "H", H, "O", O), 
    N == 0 & P != 0 ~ paste0("C", C, "H", H, "O", O, "P", P),
    TRUE ~ "formula error")
)

#create config file
config_neg <- left_join(batch4_db_adductremoved, peak_info_neg, by = c("Compound" = "compound")) %>% mutate(adduct = "M-H") %>% select(c(Compound, Identifier, mz_db, Formula_no_adduct, adduct, targetRtAxis, targetRtLeft, targetRtRight, sampleInd, C)) %>% dplyr::rename("name" = "Compound", "formula" = "Formula_no_adduct","mz" = "mz_db") %>% mutate(
  rtmin = targetRtAxis - 60,
  rtmax = targetRtAxis + 60,
  ppm = 10, 
  scale = 5, 
  height = 100, 
  snr = 10, 
  tracer_element = 'C',     # type of element
  tracer_isotope = '13C',     # type of isotope tracer
  tracer_number = round(as.numeric(C)*.5, 0)
) %>% select(-C) %>% filter(targetRtAxis != "NaN") 

#saveRDS(config_neg, "output/TarMet/neg/config_neg_4_5_23.rds")
```

## run TarMet
```{r}
measurements_neg <- run_TarMet_fromconfig(config = config_neg, rawfiles = rawfiles_neg, samples = samples_neg, rawDataset = rawDataset_neg, file_extension = "output/neg/SIP/4_23")

tracer_number_upd <- readxl::read_excel("output/TarMet/neg/SIP_neg_tracer_number_upd.xlsx") %>% select(-c(`...1`, name, Identifier, tracer_number))

config_neg_upd <- cbind (config_neg, tracer_number_upd) %>% select(-tracer_number) %>% dplyr::rename("tracer_number" = "tracer_number_upd") %>% filter(tracer_number > 2)

measurements_neg_upd <- run_TarMet_fromconfig(config = config_neg_upd, rawfiles = rawfiles_neg, samples = samples_neg, rawDataset = rawDataset_neg, file_extension = "output/neg/SIP/4_23")
```

## remove data for samples where monoisotopic peak is not detected
```{r}
measurements_neg_upd_filt_1 <- measurements_neg_upd %>% rownames_to_column("samples") %>% separate(samples, c("Identifier", "Channel")) %>% group_by(Identifier) %>% rowwise() %>% mutate_at(vars(starts_with("EC")), funs((case_when(Channel == 0 & . < 10000 ~ 0, TRUE ~ .)))) %>% mutate(samples = paste0(Identifier, "_", Channel)) %>% ungroup() %>% select(-c(Identifier, Channel)) %>% column_to_rownames("samples")
```

## prepare for IsoCorrectoR
```{r}
#measurement file
measurements_neg_file <- measurements_neg_upd_filt_1 %>% rownames_to_column("Measurements/Samples")
write.csv(x = measurements_neg_file, file = "output/neg/measurements_file_4_23.csv", quote = FALSE, row.names = FALSE, na = "")

#molecule file
#make molecules file (will need to subset)
molecules_neg <- batch4_db %>% select(c(Identifier, Formula, C)) %>% dplyr::rename(Molecule = Identifier) %>% mutate(Formula = paste0(Formula, "LabC", C)) %>% select(Molecule, Formula) %>% mutate(MS = "") %>%
  setNames(c("Molecule", "MS ion or MS/MS product ion", "MS/MS neutral loss")) 
write.csv(x = molecules_neg, file = "output/neg/molecule_file_4_23.csv", quote = FALSE, row.names = FALSE, na = "")

#make key for molecules file
key_neg <- batch4_db %>% select(Identifier, Compound)
#write.csv(x = key_neg, file = "output/neg/key_neg_4_23.csv", quote = FALSE, row.names = FALSE, na = "")
```

## Process data through IsoCorrectoR
```{r}
IsoCorrection(MeasurementFile = "output/TarMet/neg/measurements_file_4_23.csv", 
              ElementFile = "output/TarMet/neg/element_file.csv", 
              MoleculeFile = "output/TarMet/neg/molecule_file_4_23.csv",
              CorrectTracerImpurity = TRUE, CorrectTracerElementCore = TRUE, 
              CalculateMeanEnrichment = TRUE, UltraHighRes = FALSE, 
              DirOut = 'output/IsoCorrectoR/neg', 
              FileOut = 'result', 
              FileOutFormat = 'csv', 
              ReturnResultsObject = TRUE, CorrectAlsoMonoisotopic = FALSE, 
              CalculationThreshold = 10^-10, 
              CalculationThreshold_UHR = 8, 
              verbose = TRUE, Testmode = FALSE)
```

### Calculate paired rep (labeled vs unlabeled) relative mean enrichment & implement paired t-test with Cohen's D 
```{r}
#read in data from IsoCorrectoR
isoR_enrich_neg <- read.csv("output/neg/IsoCorrectoR/neg/2023-05-02_183514/IsoCorrectoR_result_MeanEnrichment.csv", header = TRUE, sep = ",", quote = "\"", na.strings = "NA")
#Manipulate data into long format
isoR_enrich_long_neg <- tidyr::pivot_longer(data = isoR_enrich_neg, cols = starts_with("EC"), names_to = "sample_name", values_to = "mean.enrichment") %>% dplyr::rename(Identifier = X)
#join with metadata
isoR_enrich_long_neg <- left_join(isoR_enrich_long_neg, meta_samp_neg) %>%
#join with key to get compound names
left_join(key_neg) %>%
#add column "Label" to indicate labeling treatment (unlabeled vs labeled) and "Rep" with the rep #
mutate(Label = ifelse(grepl("L", label_group), "labeled", "unlabeled"), Rep = readr::parse_number(as.character(block)), Label = as.factor(Label))
saveRDS(isoR_enrich_long_neg, "output/isoR_enrich_long_neg_4_23.RDS")

#need to filter out reps that don't have compound detected in labeled and unlabeled as well as compounds that only show up in one paired rep
isoR_enrich_long_neg_filt <- isoR_enrich_long_neg %>% group_by(group, Identifier, Compound, Rep) %>% mutate(n = n()) %>% filter(n == 2) %>% ungroup() %>% group_by(group, Identifier, Compound) %>% mutate(n2 = n()) %>% filter(n2 > 2) %>% ungroup() %>% select(c(group, Identifier, Compound, mean.enrichment, Label, Rep))

#paired t-test
t_test_neg <- isoR_enrich_long_neg_filt %>%
  group_by(group, Identifier, Compound) %>%
  t_test(mean.enrichment ~ Label, paired = TRUE, detailed = TRUE) %>%
  add_significance() 
  
#calculate effect size (Cohen's D)
cohens_d_neg <- isoR_enrich_long_neg_filt %>% 
  group_by(group, Identifier, Compound) %>%
  cohens_d(mean.enrichment ~ Label, paired = TRUE)

#calculate relative enrichment between unlabeled and labeled reps
rel_enrich_neg <- isoR_enrich_long_neg_filt %>%
  group_by(group, Identifier, Compound, Rep) %>% 
  pivot_wider(names_from = Label, values_from = mean.enrichment) %>% 
  summarise(labeled = dplyr::first(na.omit(labeled)),
      unlabeled = dplyr::first(na.omit(unlabeled))) %>%
  mutate(rel_enrichment = labeled - unlabeled) %>%
#calculate mean + sd enrichment difference between unlabeled and labeled rep pairs
  group_by(group, Identifier, Compound) %>% 
  mutate(mean_rel_enrichment = mean(rel_enrichment), 
         sd_rel_enrichment = sd(rel_enrichment), Rep = Rep)

#combine
neg_SIP <- left_join(t_test_neg, cohens_d_neg) %>% left_join(rel_enrich_neg)
saveRDS(neg_SIP, "output/neg_SIP_4_23.RDS" )
```


```{r}
combined_SIP <- rbind(pos_SIP, neg_SIP)
saveRDS(combined_SIP, "output/combined_SIP_4_23.RDS")

combined_rel_enrich <- rbind(rel_enrich_pos, rel_enrich_neg)
saveRDS(combined_rel_enrich, "output/combined_rel_enrich_4_23.RDS" )
```


```{r}
uniq_sig_intact <- rbind(t_test_pos, t_test_neg) %>% filter(p.signif != "ns" & estimate > 0) %>% select(Compound) %>% distinct()
```

