---
title: "xcms to pmartR"
author: "Kaitlin Rempfert"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged 
    number_sections: yes 
    css: stylesheet.css 
    toc: yes 
    toc_float: true 
    toc_depth: 3 
    code_folding: show 
editor_options:
  chunk_output_type: console
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# load required packages
library(tidyverse)
library(ggplot2)
library(pmartR)
library(qcrlsc)
library(MSnbase)
library(xcms)
library(sqldf)

source("scripts/XIC.R")
```


# Positive Mode

## Import positive mode filled peaks 
```{r}
dda_data_pos_fill <- readRDS("output/xcms/dda_data_pos_batch3_fill.rds")
```

## Extract positive mode peaks 
```{r}
#extract feature definitions and values from xcms object
pos_featureDefinitions <- as.data.frame(featureDefinitions(dda_data_pos_fill)) %>% rownames_to_column("Feature")
pos_featureValues <- as.data.frame(featureValues(dda_data_pos_fill, value = "maxo")) %>% rownames_to_column("Feature")
#join
pos_features <- left_join(pos_featureDefinitions, pos_featureValues)
#reformat data
data_pos <- pos_features %>% mutate(
  rt_min = as.numeric(rtmed)/60,
  group_id = Feature,
  mz = mzmed
)
```


## Look at intensity of QC peaks- constant throughout run?
```{r}
#QC on picked peaks- is intensity normalized?
pos_features_qc <- pos_featureValues %>% select(contains("EC_QC")) %>% mutate_if(is.numeric , replace_na, replace = 1)
boxplot(pos_features_qc, log="y", las=2, col = "cornflowerblue", ylab = "intensity", main = "QC Peak Intensities")
```


## Match mz and RT of features to LIQUID output
```{r}
#select just relevent columns 
data_pos <- data_pos %>% select(c(group_id, mz, rt_min, contains("EC")))

#import database
db <- read.csv("output/LIQUID/pos_LIQUID_db.csv") %>% select(-X)

#match to database
id <- sqldf("select * from data_pos data1 left join db db1 on db1.mz_db = (select db2.mz_db from data_pos data2 left join db db2 where data2.group_id = data1.group_id and abs(data2.mz - db2.mz_db) < .008 order by abs(data2.mz - db2.mz_db) asc limit 1)")
#filter to closest matches
#pos_annotated <- id %>% filter(!is.na(Compound)) %>% mutate(rt_diff = abs(rt_db - rt_min)) %>% group_by(Compound) %>% slice_min(rt_diff) %>% ungroup() %>% filter(rt_diff < .1) %>% select(-group_id)
pos_annotated <- id %>% filter(!is.na(Compound)) %>% mutate(rt_diff = abs(rt_db - rt_min)) %>% filter(rt_diff < 1) %>% rowwise() %>%
# mutate column for average of QC to select best "peak" and slice at max intensity
mutate(
  #QC_avg = mean(c_across(contains("QC"))), 
  tally = sum(!is.na(c_across((starts_with("EC")))))
       ) %>% group_by(Compound) %>% slice_max(tally) %>% slice_min(rt_diff) %>% ungroup() %>% select(-group_id)
```


## Create Inputs for pmartR
```{r}
#create lipid edata 
lipid_edata_pos <- pos_annotated %>% select(Compound, starts_with("EC")) 

#create lipid fdata 
#export sample names for fdata
fdata_pos <- pos_annotated %>% select(c(starts_with("EC"))) %>% t() %>% as.data.frame() %>% rownames_to_column("sample_name") %>% select(sample_name)
#read in metadata
meta <- readxl::read_xlsx("data/incubation_metadata_w_qc_BTLE_set2.xlsx")
#create lipid fdata (crop, timepoint, site, group)
lipid_fdata_pos <- left_join(fdata_pos, meta) %>% select(c(sample_name, group, batch, injection_order, site, crop, timepoint)) 

#create lipid emeta
lipid_emeta_pos <- pos_annotated %>% select(Compound, Category, Adduct, Formula, mz_db, rt_db)
```

## Create lipidomic data object in pmartR
```{r}
mylipidData_pos <- pmartR::as.lipidData(e_data = lipid_edata_pos, f_data = lipid_fdata_pos, e_meta = lipid_emeta_pos, edata_cname = "Compound", emeta_cname = "Compound", fdata_cname = "sample_name", data_scale = "abundance")
class(mylipidData_pos)

saveRDS(mylipidData_pos, "output/pmartR/mylipidData_pos_batch3.rds")

plot(mylipidData_pos)
```



# Negative Mode

## Import positive mode filled peaks 
```{r}
dda_data_neg_fill <- readRDS("output/xcms/dda_data_neg_batch4_fill.rds")
```

## Extract positive mode peaks 
```{r}
#extract feature definitions and values from xcms object
neg_featureDefinitions <- as.data.frame(featureDefinitions(dda_data_neg_fill)) %>% rownames_to_column("Feature")
neg_featureValues <- as.data.frame(featureValues(dda_data_neg_fill, value = "maxo")) %>% rownames_to_column("Feature")
#join
neg_features <- left_join(neg_featureDefinitions, neg_featureValues)
#reformat data
data_neg <- neg_features %>% mutate(
  rt_min = as.numeric(rtmed)/60,
  group_id = Feature,
  mz = mzmed
)
```


## Look at intensity of QC peaks- constant throughout run?
```{r}
#QC on picked peaks- is intensity normalized?
neg_features_qc <- neg_featureValues %>% select(contains("EC_QC")) %>% mutate_if(is.numeric , replace_na, replace = 1)
boxplot(neg_features_qc, log="y", las=2, col = "cornflowerblue", ylab = "intensity", main = "QC Peak Intensities")
```


## Match mz and RT of features to LIQUID output
```{r}
#select just relevent columns 
data_neg <- data_neg %>% select(c(group_id, mz, rt_min, contains("EC")))

#import database
db <- read.csv("data/databases/neg_LIQUID_db.csv") %>% select(-X)

#match to database
id <- sqldf("select * from data_neg data1 left join db db1 on db1.mz_db = (select db2.mz_db from data_neg data2 left join db db2 where data2.group_id = data1.group_id and abs(data2.mz - db2.mz_db) < .008 order by abs(data2.mz - db2.mz_db) asc)")

# removed "limit 1" after asc
#filter to closest matches
neg_annotated <- id %>% filter(!is.na(Compound)) %>% mutate(rt_diff = abs(rt_db - rt_min)) %>% filter(rt_diff < 1) %>% rowwise() %>%
# mutate column for average of QC to select best "peak" and slice at max intensity
mutate(
  #QC_avg = mean(c_across(contains("QC"))), 
  tally = sum(!is.na(c_across((starts_with("EC")))))
       ) %>% group_by(Compound) %>% slice_max(tally) %>% slice_min(rt_diff) %>% ungroup() %>% select(-c(group_id, tally))
```


## Manually extract fatty acid peaks (shift in RT between samples)
```{r}
#import FA database
FA <- readxl::read_xlsx("data/databases/fatty_acids.xlsx") %>% mutate(rt_window = 80, Adduct = "[M-H]-")

#extract peaks corresponding to fatty acids 
peak_area <- extract_peak_apex_h(FA, dda_data_neg_fill)

#reformat peak area data (filtered out QC here)
peak_area_reformatted <- peak_area %>% ungroup() %>% group_by(Compound) %>% mutate(
  mz = as.numeric(as.character(mz)), 
  rt_min = as.numeric(as.character(rt_min)),
  area = as.numeric(as.character(area)),
  rt_min = mean(rt_min, na.rm = TRUE), 
  mz = mean(mz, na.rm = TRUE)) %>% ungroup() %>% pivot_wider(names_from = sample_name, values_from = area)

#join with peak area metadata and rename columns for joining with other annotated data
FA_peakareas <- left_join(peak_area_reformatted, FA, by = c("Compound" = "compound")) %>% select(-rt_window) %>% mutate(
  Category = "Fatty Acyls [FA]", 
  Adduct = "[M-H]-",
  rt_diff = abs(mz_db - mz)
)

#join with other annotated data
neg_annotated <- rbind(neg_annotated, FA_peakareas)
```


## Create Inputs for pmartR
```{r}
#create lipid edata 
lipid_edata_neg <- neg_annotated %>% select(Compound, starts_with("EC")) 

#create lipid fdata 
#export sample names for fdata
fdata_neg <- lipid_edata_neg %>% select(starts_with("EC")) %>% t() %>% as.data.frame() %>% rownames_to_column("sample_name") %>% select(sample_name)
#read in metadata
meta <- readxl::read_xlsx("data/incubation_metadata_w_qc_BTLE_set2.xlsx")
#create lipid fdata (crop, timepoint, site, group)
lipid_fdata_neg <- left_join(fdata_neg, meta) %>% select(c(sample_name, group, batch, injection_order, site, crop, timepoint)) 

#create lipid emeta
lipid_emeta_neg <- neg_annotated %>% select(Compound, Category, Adduct, Formula, mz_db, rt_db)
```

## Create lipidomic data object in pmartR
```{r}
mylipidData_neg <- pmartR::as.lipidData(e_data = lipid_edata_neg, f_data = lipid_fdata_neg, e_meta = lipid_emeta_neg, edata_cname = "Compound", emeta_cname = "Compound", fdata_cname = "sample_name", data_scale = "abundance")
class(mylipidData_neg)

saveRDS(mylipidData_neg, "output/pmartR/mylipidData_neg_batch4.rds")

plot(mylipidData_neg)
```

